---
title: "Generating ELM/CLM domain and surface files for a single point"
author: "Marcos Longo"
date: "16-Jun-2021"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

# Introduction

This document shows how to generate domain and surface files for a single point of interest.
This is needed for when setting a new site. Currently this script simply uses existing domain 
and surface files, extracts information from the nearest grid cell, and complements with 
site-specific information you might have.

# Reset session:

Use this chunk to fully reset R.
```{r echo=FALSE, message=FALSE, results='hide'}
# Unload all packages except for the R default ones
plist = names(sessionInfo()$otherPkgs)
if (length(plist) > 0){
   dummy = sapply(X=paste0("package:",plist),FUN=detach,character.only=TRUE,unload=TRUE)
}#end if (length(plist) > 0)


# Remove all variables
rm(list=ls())

# Reset warnings
options(warn=0)

# Close all plots
invisible(graphics.off())

# Clean up
invisible(gc())
```

# Main settings

## Required packages

Load some packages. By default, we skip the warning messages, but it may be a good idea to
visualise them if it is the first time running.
```{r, label='load-packages', message=FALSE, results='hide'}

isfine = c( data.table = require(data.table,quietly=TRUE,warn.conflicts=FALSE)
          , fields     = require(fields    ,quietly=TRUE,warn.conflicts=FALSE)
          , lubridate  = require(lubridate ,quietly=TRUE,warn.conflicts=FALSE)
          , ncdf4      = require(ncdf4     ,quietly=TRUE,warn.conflicts=FALSE)
          , purrr      = require(purrr     ,quietly=TRUE,warn.conflicts=FALSE)
          , reshape2   = require(reshape2  ,quietly=TRUE,warn.conflicts=FALSE)
          , tidyverse  = require(tidyverse ,quietly=TRUE,warn.conflicts=FALSE)
          )#end c
if (! all(isfine)){
   cat0(" List of required packages, and the success status loading them:\n")
   print(isfine)
   stop(" Some packages are missing and must be installed.")
}
```


## Path and file location

Set paths and files for input and output.

* **here_path**. The main path for processing data.
* **input_path**. The location where existing surface and domain netcdf files are located
* **output_path**. Directory to where FATES files will be written.

```{r, label='path-settings'}
here_path   = getwd()
input_path  = file.path(here_path,"input_netcdf")
output_path = file.path(here_path,"fates_domain+surface_files")
```

Set the URL for the input surface data. We will only download the data in case we cannot locate these files at `input_netcdf`.
If you have your own set of input data, set a dummy URL with the actual file name.  For example, if you have the domain file `my_surface_data_file.nc`,
 set `url.domain="https://localhost/my_surface_data_file.nc"`.
```{r, label='input-url'}
url_surfdata = "https://svn-ccsm-inputdata.cgd.ucar.edu/trunk/inputdata/lnd/clm2/surfdata_map/release-clm5.0.18/surfdata_0.9x1.25_hist_78pfts_CMIP6_simyr2000_c190214.nc"
```


## Author information
Provide information about yourself. This will be included in the netCDF headers. 
In case you do not want to specify any of them, set them to `NA_character_`

* **author_name**. Person who is generating these files (typically it's you).
* **author_email**. Contact email (so people can reach out to you if problems arise).
```{r }
author_name  = "Marcos Longo"
author_email = "mlongo@lbl.gov"
```

## Site information

Provide some basic information about the site:

* **xid**. Unique site ID.
* **site_desc**. Site description (to be added to the NetCDF files headers).
* **site_lon**. Longitude (degrees east).
* **site_lat**. Latitude (degrees north).
* **site_alt**. Altitude (metres above sea level).
* **dxy**. Intended grid size.
* **site_pcsand**; **site_pcclay**. Percentage of sand and clay, respectively.  They should be both be limited between 0 and 100., and their sum shall not exceed 100.  Silt fraction will be derived from these fractions.  If these fractions are not known, set them to `NA`.  If texture information is provided, we assume that all layers have
the same texture (this should be revisited once I figure out the soil layers).
* **site_organic**. Soil organic content $\left[\mathrm{kg\,m^{-3}}\right]$.  For now we use this value to rescale the gridded organic content of all input layers.
* **site_bedrock**. Soil depth to bedrock $\left[\mathrm{m}\right]$. In case depth is unknown, leave it as `NA_real_`.
* **undef**. A number to denote missing values. This is needed by NetCDF, but the data should not have any.
* **force_natural**. Should we overwrite the fraction cover of global gridded data set, so natural vegetation becomes 100%?

```{r, label='site-settings'}
xid           = "gyf"
site_desc     = "Guyaflux tower at Paracou, French Guiana"
dat_version   = "1.0"
site_lon      = -52.912222
site_lat      =   5.281667
site_alt      =  30.000000
dxy           =   0.1
site_xsand    = 56.2
site_xclay    = 37.2
site_organic  = 31.1
site_bedrock  = NA_real_
undef         = -9999.99
force_natural = FALSE
```


Additional settings for retrieving the surface gridded data

* **pcveg_min**. Minimum percentage of natural vegetation to consider when loading surface grid. This can be useful to avoid selecting a grid cell over the ocean when using a coastal site.
```{r, label='surf-data-settings'}
pcveg_min = 8.00
```

This concludes the initial settings.  From this point on, you may not need to change anything,
unless you are debugging or adding new features.


# Data processing

## Check inputs

Ensure the input and output paths
```{r, label='path-checks', message=FALSE, results='hide'}
dummy = dir.create(input_path ,showWarnings=FALSE,recursive=TRUE)
dummy = dir.create(output_path,showWarnings=FALSE,recursive=TRUE)
```

First, we check whether the input surface data file is already available at the input path. In case it is not, we 
download it.
```{r, label='load-inputs'}

# Find base name
input_surfdata = file.path(input_path,basename(url_surfdata))

# Download data in case they are not found.
if (! file.exists(input_surfdata)) download.file(url=url_surfdata,destfile=input_surfdata)

```


## Output names

Build the output file names, following the ELM/CLM convention.
```{r, label='set-output-names'}
# Labels for describing the data set.
ymd_now = today(tzone = "UTC")
ymd_lab = sprintf("%4.4i%2.2i%2.2i",year(ymd_now),month(ymd_now),day(ymd_now))
dxy_lab = sprintf("%.1fx%.1f",dxy,dxy)

# Suffix for files.
suffix  = paste0(xid,"_",dxy_lab,"_v",dat_version,"_c",ymd_lab)

# Domain file.  In case the file exists, remove it.
output_domain   = file.path(output_path,paste0("domain_lnd_",suffix,".nc"))
if (file.exists(output_domain)) file.remove(output_domain)

# Surface data file
output_surfdata = file.path(output_path,paste0("surfdata_",suffix,".nc"))
if (file.exists(output_surfdata)) file.remove(output_surfdata)
```

## Domain file

The domain file is written entirely from the information provided at the beginning of this document.

First, we set the centre and edge of the grid cells. Make sure longitude is always set from $0^\circ$ to $360^\circ$, and that the centre of the latitude points are not too close to the poles.
```{r, label='set-coords'}
# Ensure the longitude and latitude are correctly bounded.
clon = site_lon %% 360.
clat = max(-90+0.5*dxy,min(90-0.5*dxy,site_lat))

# Find edges
wlon = (clon - 0.5 * dxy) %% 360.
elon = (clon + 0.5 * dxy) %% 360.
slat = max(-90.,clat - 0.5 * dxy)
nlat = min(+90.,clat + 0.5 * dxy)
```


Find the area of this grid cell, in steradians, using the definition of solid angle $\omega$:
$$\begin{align}
& \omega = \left[\sin{\left(\varphi_{N}\right)} - \sin{\left(\varphi_{S}\right)} \right] \, \left[ \left(\lambda_{E} - \lambda_{W} \right) \mod{2\,\pi{}} \right]
\end{align}$$
where $\varphi_{S}$ and $\varphi_{N}$ are the latitudes of the southern and northern edges [$\mathrm{rad}$], and $\lambda_{W}$ and $\lambda_{E}$ are the longitude of the western and eastern edges [$\mathrm{rad}$]. The modulo operator ensures that the results are correct when the centre of the grid cell crosses the Greenwich meridian.
```{r, label='solid-angle'}
# Turn angles to radians.
wlonr = wlon * pi / 180.
elonr = elon * pi / 180.
slatr = slat * pi / 180.
nlatr = nlat * pi / 180.

# Find solid angle area (assuming radius=1 so results are in steradians).
solid_angle = (sin(nlatr)-sin(slatr)) * ((elonr-wlonr) %% (2*pi))
```

Define the dimensions for the new file. It should be always a single point:
```{r, label='domain-dims'}
# Dimensions are unitless
ni = ncdim_def( name="ni",units="",vals=sequence(1L),create_dimvar=FALSE)
nj = ncdim_def( name="nj",units="",vals=sequence(1L),create_dimvar=FALSE)
nv = ncdim_def( name="nv",units="",vals=sequence(4L),create_dimvar=FALSE)

# List of dimensions
ij  = list(ni,nj)
vij = list(nv,ni,nj)
```

Create a list with the variables to be added to the single-point domain file.  Each list element 
contains the following variables.

* **vals**. The values to be included in the variables.
* **long**. The variable long (and descriptive) name.
* **unit**. The variable units. In case the variable does not have units, leave it blank (`""`).
* **dims**. The variable dimensions (a list with the dimensions defined above).
* **attx**. Other relevant information to be added as attributes. In case no additional attribute is needed, set this to `list()`
```{r, label='set-domain'}
# List of data to write.
nco_dat      = list()
nco_dat$xc   = list( vals = array(data = clon                  , dim =   c(1,1))
                   , long = "longitude of the grid cell centre"
                   , unit = "degrees_east"
                   , dims = ij
                   , attx = list( bounds = "xv")
                   )#end list
nco_dat$yc   = list( vals = array(data = clat                  , dim =   c(1,1))
                   , long = "latitude of the grid cell centre"
                   , unit = "degrees_north"
                   , dims = ij
                   , attx = list( bounds = "yv")
                   )#end list
nco_dat$xv   = list( vals = array(data = c(wlon,elon,elon,wlon), dim = c(4,1,1))
                   , long = "longitude of grid cell vertices"
                   , unit = "degrees_east"
                   , dims = vij
                   , attx = list(note = "Vertex order: 1 - SW; 2 - SE; 3 - NE; 4 - NW")
                   )#end list
nco_dat$yv   = list( vals = array(data = c(slat,slat,nlat,nlat), dim = c(4,1,1))
                   , long = "latitude of grid cell vertices"
                   , unit = "degrees_north"
                   , dims = vij
                   , attx = list(note = "Vertex order: 1 - SW; 2 - SE; 3 - NE; 4 - NW")
                   )#end list
nco_dat$mask = list( vals = array(data =          1L           , dim =   c(1,1))
                   , long = "land domain mask"
                   , unit = ""
                   , dims = ij
                   , attx = list( note        = "Unitless."
                                , coordinates = "xc yc"
                                , classes     = "0 - water (not active); 1 - land (active)"
                                )#end list
                   )#end list
nco_dat$frac = list( vals = array(data =          1.           , dim =   c(1,1))
                   , long = "fraction of grid cell that is active"
                   , unit = ""
                   , dims = ij
                   , attx = list( note        = "Unitless."
                                , coordinates = "xc yc"
                                , filter1     = paste0( "error if frac> 1.0+eps or "
                                                      , "frac < 0.0-eps; eps = 0.1000000E-11"
                                                      )#end paste0
                                , filter2     = paste0( "limit frac to [fminval,fmaxval];"
                                                      , "fminval= 0.1000000E-02 fmaxval=  1.000000"
                                                      )#end paste0
                                )#end list
                   )#end list
nco_dat$area = list( vals = array(data = solid_angle           , dim =   c(1,1))
                   , long = "area (solid angle) of grid cell, in steradian"
                   , unit = "sr"
                   , dims = ij
                   , attx = list( coordinates = "xc yc")
                   )#end list
```


Define the netCDF variables for the output variables.  
```{r, label='def-domain-vars'}
nco_vlist = list()
for (v in seq_along(nco_dat)){
   # Load information for this variable.
   v_vnow   = names(nco_dat)[v]       # Variable name
   v_long   = nco_dat[[v_vnow]]$long  # Variable description ("long name")
   v_unit   = nco_dat[[v_vnow]]$unit  # Units
   v_dims   = nco_dat[[v_vnow]]$dims  # Variable dimensions

   # Define netCDF variable
   nco_vlist[[v_vnow]] = ncvar_def( name     = v_vnow
                                  , units    = v_unit
                                  , dim      = v_dims
                                  , longname = v_long
                                  )#end ncvar_def
}#end for (v in seq_along(nci_vlist))
```


Create the output netCDF file.
```{r, label='gen-nc4-domain'}

# Create the netCDF connection
if ("nc_output" %in% ls()){ dummy=nc_close(nc_output); file.remove(output_domain); rm(nc_output)}
nc_output = nc_create(filename=output_domain,vars=nco_vlist)

# Define the code developer information (indirect way so the email is not visible).
developer_name  = c( 111L, 103L, 110L, 111L,  76L,  32L, 115L, 111L,  99L, 114L,  97L,  77L)
developer_email = c( 118L, 111L, 103L,  46L, 108L,  98L, 108L,  64L, 111L, 103L, 110L, 111L
                   , 108L, 109L)

# Generate information about this file.
glob_att = list( title           = paste0( "Domain file for ",site_desc,".")
               , version         = dat_version
               , date_created    = paste0(as.character(now(tzone="UTC"))," UTC")
               , source_code     = "make_fates_domain+surface.Rmd"
               , code_notes      = "Domain file compatible with ELM-FATES and CLM-FATES"
               , code_developer  = paste0( intToUtf8(rev(developer_name))
                                         ," <"
                                         , intToUtf8(rev(developer_email))
                                         ,">"
                                         )#end paste0
               , file_author     = paste0(author_name," <",author_email,">")
               )#end list

# Add information to the global attribute list
for (l in seq_along(glob_att)){
   # Current attribute information
   att_name  = names(glob_att)[l]
   att_value = glob_att[[l]]

   # Add attribute 
   dummy = ncatt_put(nc=nc_output,varid=0,attname=att_name,attval=att_value)
}#end for (l in seq_along(global_att))
```

Populate the netCDF with the domain variables, and include the attributes for each variable.
```{r, label='add-var-domain'}
for (v in seq_along(nco_dat)){
   # Retrieve variable information
   v_vnow  = names(nco_dat)[v]
   nco_now = nco_dat[[v_vnow]]

   # Add variable values
   dummy = ncvar_put(nc = nc_output,varid=v_vnow,vals=nco_now$vals)

   # Add extra attributes
   for (a in seq_along(nco_now$attx)){
      # Retrieve attribute information
      a_name  = names(nco_now$attx)[a]
      a_value = nco_now$attx[[a]] 
      
      dummy = ncatt_put(nc = nc_output, varid=v_vnow,attname=a_name,attval=a_value)
   }#end for (a in seq_along(v_attx))
}#end for (v in seq_along(nco_vlist))
```


Save new netCDF, and close the output file. We remove the connection from memory.
```{r, label='end-var-domain'}
   dummy = nc_close(nc_output)
   rm(nc_output)
```

# Surface data file

To generate the grid surface data file, we will load an existing surface data file, locate the grid cell
that is the closest to the point of interest, abd extract the data.  For some of the variables, such as 
soil texture and soil organic carbon, there may be local observations.  If this is the case, we replace the gridded information with the local observations.

## Load the gridded surface data

Open the surface data file, and retrieve the grid dimensions.
```{r, label='load-surfdata-gridded'}
# Open gridded surface data
if ("nc_input" %in% ls()){dummy=nc_close(input_surfdata);rm(nc_input)}
nc_input = nc_open(filename=input_surfdata)
```


Retrieve global attributes from the input file, and add information about this processing.
```{r, label='set-surfdata-glob-att'}

# Retrieve original attributes. We only make minor edits.
glob_att = ncatt_get(nc = nc_input, varid = 0)


# Define the code developer information (indirect way so the email is not visible).
developer_name  = c( 111L, 103L, 110L, 111L,  76L,  32L, 115L, 111L,  99L, 114L,  97L,  77L)
developer_email = c( 118L, 111L, 103L,  46L, 108L,  98L, 108L,  64L, 111L, 103L, 110L, 111L
                   , 108L, 109L)

# Get current time for the log
create_time = paste0(as.character(now(tzone="UTC"))," UTC")

# Generate information about this file.
glob_new = list( title           = paste0( "Domain file for ",site_desc,".")
               , Version         = dat_version
               , Revision_Id     = paste0( "$Id: make_fates_domain+surface.Rmd"
                                         , " 0 ",create_time,Sys.info()["user"],"$"
                                         )#end paste0
               , History_Log     = paste0( "created on: ", create_time)
               , Logname         = paste0( "code originally developed by "
                                         , intToUtf8(rev(developer_name))
                                         ," <"
                                         , intToUtf8(rev(developer_email))
                                         ,">"
                                         )#end paste0
               , Host            = Sys.info()["nodename"]
               , File_Author     = paste0(author_name," <",author_email,">")
               )#end list

# Update attribute list
glob_att = modifyList( glob_att, glob_new)
```



Load coordinates, and identify the nearest grid point. 
```{r label='find-near-neighbour'}
# Retrieve longitude and latitude
in_lon   = melt(ncvar_get(nc=nc_input,"LONGXY"    )); names(in_lon) = c("x","y","lon"  )
in_lat   = melt(ncvar_get(nc=nc_input,"LATIXY"    )); names(in_lat) = c("x","y","lat"  )
in_veg   = melt(ncvar_get(nc=nc_input,"PCT_NATVEG")); names(in_veg) = c("x","y","pcveg")

# Create a tibble with coordinates
in_coord = tibble(merge(merge(in_lon,in_lat),in_veg)) 


# Define function that takes longitude/latitude and find distances.
rdist_wrap = function(lon,lat,lon0,lat0){
   ans = c(fields::rdist.earth(x1=cbind(lon0,lat0),x2=cbind(lon,lat),miles=FALSE))
   return(ans)
}#end function

# Find the distances.
in_coord = in_coord %>%
           mutate( dist = rdist_wrap(lon=lon,lat=lat,lon0=site_lon,lat0=site_lat))


# Select the nearest neighbour
in_near = in_coord %>% filter(pcveg > pcveg_min) %>% slice(which.min(dist))
in_xy   = in_near %>% select(c(x,y)) %>% unlist(); names(in_xy) = NULL
```

Load all the dimensions from the input netCDF file, and create dimensions for the output
file that are essentially copies of the input. The only exceptions are the longitude and
latitude, which will be always one (single site).
```{r label='set-surf-dimensions'}

# List of dimension names
nc_dnames = names(nc_input$dim)

# Loop through all the dimensions
nc_dims = list()
for (d in seq_along(nc_dnames)){
   # Retrieve dimensions
   d_dname   = nc_dnames[d]
   nc_dimnow = nc_input$dim[[d_dname]]

   # Fix dimensions and values for longitude and latitude.
   if (d_dname %in% c("lsmlon","lsmlat")){
      # Longitude or latitude, make data a single point
      nc_dimnow = modifyList( x   = nc_dimnow
                            , val = list( len = 1L, vals = sequence(1L))
                            )#end modifyList
   }#end if (nc_dimnow)

   # Decide whether to append calendar information
   if (nc_dimnow$unlim && (! is.null(nc_dimnow$calendar))){
      nc_calendar = nc_dimnow$calendar
   }else{
      nc_calendar = NA_character_
   }#end if (nc_dimnow$unlim && (! is.null(nc_dimnow$calendar)))

      
   # Create dimensions. Distinguish unlimited and limited dimensions (if needed).
   nc_dims[[d_dname]] = ncdim_def( name          = d_dname
                                 , units         = nc_dimnow$units
                                 , vals          = nc_dimnow$vals
                                 , unlim         = nc_dimnow$unlim
                                 , create_dimvar = nc_dimnow$create_dimvar
                                 , calendar      = nc_calendar
                                 )#end ncdim_def
}#end for (d in seq_along(nc_dnames))
```


Define variables that should feature in the output. Every variable in the input should be in the output too.
```{r, label='set-surf-variables'}

# List of variable names
nc_vnames = names(nc_input$var)

# Loop through all the variables, retrieve data and
nc_vars   = list()
nc_values = list()
for (v in seq_along(nc_vnames)){
   # Retrieve variable information
   v_vname = nc_vnames[v]
   v_vinfo = nc_input$var [[v_vname]]
   v_vlong = v_vinfo$longname
   v_vunit = v_vinfo$unit
   v_vmiss = v_vinfo$missval

   # Switch the longitude and latitude dimensions with point
   if (is.null(v_vinfo$varsize)){
      # Dummy dimensions
      v_count = NA
      v_start = NA
      v_vdims = list()
   }else{
      # Load count, and assume initially that all dimensions start at 1.
      v_count = v_vinfo$varsize
      v_start = rep(1L,times=length(v_count))

      # Find dimensions 
      v_dname = c(unlist(do.call(rbind,v_vinfo$dim)[,"name"]))
      v_vdims = nc_dims[v_dname]
      
      # Find longitude and latitude dimensions
      ixy  = match(c("lsmlon","lsmlat"),v_dname)
      sel  = ! is.na(ixy)

      # Make sure we only read the data from the nearest neighbour
      v_count[ixy[sel]]   = c(1,1)[sel]
      v_start[ixy[sel]]   = in_xy [sel]
      names(v_start) = NULL
   }#end if (length(v_start) == 0)

   # Load information to define the data set.
   nc_vars[[v_vname]] = ncvar_def( name     = v_vname
                                 , units    = v_vunit
                                 , dim      = v_vdims
                                 , missval  = v_vmiss
                                 , longname = v_vlong
                                 )#end ncvar_def

   # Load the data 
   nc_values[[v_vname]] = ncvar_get( nc             = nc_input
                                   , varid          = v_vname
                                   , start          = v_start
                                   , count          = v_count
                                   , collapse_degen = FALSE
                                   )#end ncvar_get
}#end for (v in seq_along(nc_vnames))
```


In case texture and soil organic content were provided, we use the site-specific information instead. Currently the site-specific soil texture will be assumed constant, and the site-specific soil organic contant will be used to re-scale all the layers.  We also replace the longitude and latitude information with the site-specific information. Finally, in case `force_natural` is true, we impose percentage of natural vegetation to 100%.
```{r, label='set-site-specific'}
# Re-set coordinates
nc_values$LONGXY = 0. * nc_values$LONGXY + site_lon
nc_values$LATIXY = 0. * nc_values$LATIXY + site_lon

# Re-define percentage of natural vegetation to 100%, and other classes to 0%
if (force_natural){
   nc_values$PCT_NATVEG  = 0. * nc_values$PCT_NATVEG + 100.
   nc_values$PCT_URBAN   = 0. * nc_values$PCT_URBAN
   nc_values$PCT_CROP    = 0. * nc_values$PCT_CROP
   nc_values$PCT_WETLAND = 0. * nc_values$PCT_WETLAND
   nc_values$PCT_LAKE    = 0. * nc_values$PCT_LAKE
   nc_values$PCT_GLACIER = 0. * nc_values$PCT_GLACIER
}#end if (force_natural)


# In case texture was defined, set the values.
if (! ( is.na(site_xsand) || is.na(site_xclay) )){
   # Find silt fraction
   site_xsilt = 100. - site_xsand - site_xclay

   # Set a vector with textures to test whether or not to use site-specific data
   site_xtext = c(site_xsand,site_xsilt,site_xclay)

   # Make sure the texture provided is valid.
   if (all( (site_xtext >= 0.) & (site_xtext <= 100.) ) ){
      # Replace sand and clay with the site-specific data
      nc_values$PCT_SAND = 0. * nc_values$PCT_SAND + site_xsand
      nc_values$PCT_CLAY = 0. * nc_values$PCT_CLAY + site_xclay

      # Update global attribute list to inform that the data are site-specific
      map_soil_texture_file = "site-specific information"
      glob_att = modifyList( x   = glob_att
                           , val = list( map_soil_texture_file = map_soil_texture_file )
                           )#end modifyList
   }#end if (all( (site_xtext => 0.) & (site_xtext <= 100.) ))
}#end if (! ( is.na(site_xsand) || is.na(site_xclay) ))

# In case site-specific organic content was defined, rescale the gridded values.
is_site_organic = ! is.na(site_organic)
if (is_site_organic){
   f_scal = site_organic / max(c(nc_values$ORGANIC))
   nc_values$ORGANIC = f_scal * nc_values$ORGANIC   

   # Update global attribute list to inform that the data are site-specific
   map_soil_organic_file = paste0( glob_att$map_soil_texture_file
                                 ," scaled with site-specific information.")
   glob_att = modifyList( x   = glob_att
                        , val = list( map_soil_texture_file = "site-specific information" )
                        )#end modifyList
}#end if (! is.na(site_organic))
```

In case depth to bedrock in known, include it in the list of variables.
```{r, label='set-bedrock'}
if (! is.na(site_bedrock)){
   # Append depth to bedrock to surface data.
   nc_vars$zbedrock = ncvar_def( name     = "zbedrock"
                               , units    = "m"
                               , dim      = nc_dims[c("lsmlon","lsmlat")]
                               , missval  = nc_input$var$PCT_SAND$missval
                               , longname = "soil depth"
                               )#end ncvar_def 

   # Append value to the list of values
   nc_values$zbedrock = array(data=site_bedrock,dim=c(1,1))
}#end if (! is.na(site_bedrock))
```


Create the output files.
```{r, label='gen-output'}

# Create output netCDF
nc_output = nc_create(filename = output_surfdata,vars=nc_vars)

# Add global attributes
for (a in seq_along(glob_att)){
   # Retrieve attribute information
   a_name  = names(glob_att)[a]
   a_value = glob_att[[a_name]] 

   # Include global attribute      
   dummy = ncatt_put(nc = nc_output, varid=0,attname=a_name,attval=a_value)
}#end for (a in seq_along(glob_att))

# Add variables
for (v in seq_along(nc_values)){
   # Retrieve variable information
   v_name  = names(nc_values)[v]
   v_value = nc_values[[v_name]] 

   # Add variable values
   dummy = ncvar_put(nc = nc_output,varid=v_name,vals=v_value)
}#end for (v in seq_along(nc_values))

# Close input and output files
dummy = nc_close(nc = nc_input)
dummy = nc_close(nc = nc_output)
```

