---
title: "FATES met driver converter."
author: "Marcos Longo"
date: "11-Jun-2021"
output:
  html_document:
    df_print: paged
  pdf_document: default
references:
- id: bolton1980mwr
  title: The computation of equivalent potential temperature
  author: 
  - family: Bolton
    given: David
  container-title: Mon. Wea. Rev.
  URL: https://doi.org/10.1175/1520-0493(1980)108<1046:TCOEPT>2.0.CO;2
  DOI: 10.1175/1520-0493(1980)108<1046:TCOEPT>2.0.CO;2
  issue: 108
  page: 1046-1053
  issued:
     year: 1980
---

# Reset session:

Use this chunk to fully reset R.
```{r echo=FALSE, message=FALSE, results='hide'}
# Unload all packages except for the R default ones
plist = names(sessionInfo()$otherPkgs)
if (length(plist) > 0){
   dummy = sapply(X=paste0("package:",plist),FUN=detach,character.only=TRUE,unload=TRUE)
}#end if (length(plist) > 0)


# Remove all variables
rm(list=ls())

# Reset warnings
options(warn=0)

# Close all plots
invisible(graphics.off())

# Clean up
invisible(gc())
```

# Introduction

This script reads in hourly meteorological data from towers or automatic weather stations,
and generates the meteorological driver files for FATES.  The input file must be a csv 
file containing the following variables:

* Year
* Month
* Day
* Hour
* Atmospheric pressure
* Air temperature
* Humidity. One of the following must be provided (and this order of preference):
  - Specific humidity
  - Mixing ratio
  - Vapour pressure 
  - Relative humidity
* Wind speed
* Incident solar irradiance
* Precipitation

The example below shows how the data should look like (first 10 lines):

```
year,month,day,hour,min,atm.prss,atm.tmp,atm.rhv,atm.vels,atm.vdir,rshort.in,rlong.in,rain
2004,1,1,0,0,999.5,24.41,97.87,2.9,58.8,0,443.76,0
2004,1,1,1,0,999.5,24.63,96.97,2.55,65.1,0,444.67,0
2004,1,1,2,0,999.5,24.77,96.9,2.15,59.5,0,445.67,0
2004,1,1,3,0,998.5,24.94,96.56,2.45,55.4,0,446.41,0
2004,1,1,4,0,998.5,24.21,95.99,4.4,92.8,0,438.98,4.2
2004,1,1,5,0,998.5,23.73,98.16,3.5,101.8,0,437.42,0.2
2004,1,1,6,0,998.5,24.03,98.16,4.05,76.9,0,440.88,2.6
2004,1,1,7,0,999.5,23.44,98.5,4.65,64.9,1.09,436.54,2.8
2004,1,1,8,0,1000.5,23.54,98.73,1.95,96.9,33.97,433.63,0
2004,1,1,9,0,1001.5,23.93,98.9,0.75,211.2,93.71,432.94,0
2004,1,1,10,0,1002,24.67,98.28,0.3,206.8,239.25,419.64,0
```

**Important**: All input data must be gap-filled beforehand.  This script will not 
gap fill the data for you.

# Main settings

## Required packages

Load some packages. By default, we skip the warning messages, but it may be a good idea to
visualise them if it is the first time running.
```{r, label='load-packages', message=FALSE, results='hide'}

isfine = c( data.table = require(data.table,quietly=TRUE,warn.conflicts=FALSE)
          , lubridate  = require(lubridate ,quietly=TRUE,warn.conflicts=FALSE)
          , ncdf4      = require(ncdf4     ,quietly=TRUE,warn.conflicts=FALSE)
          , purrr      = require(purrr     ,quietly=TRUE,warn.conflicts=FALSE)
          , reshape2   = require(reshape2  ,quietly=TRUE,warn.conflicts=FALSE)
          , tidyverse  = require(tidyverse ,quietly=TRUE,warn.conflicts=FALSE)
          )#end c
if (! all(isfine)){
   cat0(" List of required packages, and the success status loading them:\n")
   print(isfine)
   stop(" Some packages are missing and must be installed.")
}
```

## Paths

Set paths and files for input and output.

* **here_path**. The main path for processing data.  The output path will be generated here.
* **input_file**. The csv file containing the gap-filled tower/AWS data (full path).

```{r, label='path-settings'}
here_path   = getwd()
input_file  = file.path(here_path,"input_csv","gyf_met-driver_2004-2019.csv")
```

## Author information
Provide information about yourself. This will be included in the netCDF headers. 
In case you do not want to specify any of them, set them to `NA_character_`

* **author_name**. Person who is generating these files (typically it's you).
* **author_email**. Contact email (so people can reach out to you if problems arise).
* **datprov_name**. Data provider name (who shared the met drivers with you).
* **datprov_email**. Data provider email.
* **data_usage_notes**. Brief data usage note message.
```{r }
author_name  = "Marcos Longo"
author_email = "mlongo@lbl.gov"
datprov_name = "Damien Bonal"
datprov_email = "damien.bonal@inrae.fr"
data_usage_notes = paste0( " If you plan to use these data for any scientific analysis,"
                         , " you should contact the data provider first and ask for permission"
                         , " to use the data and check with them how to acknowledge their"
                         , " contribution (including, but not limited to offer co-authorship)."
                         )#end paste0
```

## Tower information

Provide some basic information about the tower:

* **xid**. Unique Tower ID.
* **site_desc**. Site description (to be appended to the netCDF headers)
* **dat_version**. Version of this data set.
* **site_lon**. Site longitude (degrees east).
* **site_lat**. Site latitude (degrees north).
* **site_alt**. Site altitude (metres above sea level).
* **site_refhgt**. Height reference for observations (e.g., the tower height above ground).
* **dxy**. Intended grid size (we will generate a 1x1 pseudo-grid for FATES).
* **undef**. A number to denote missing values. This is needed by NetCDF, but the data should not have any

```{r, label='met-driver-settings'}
xid         = "gyf"
site_desc   = "Guyaflux tower at Paracou, French Guiana"
dat_version = "1.1"
site_lon    = -52.912222
site_lat    =   5.281667
site_alt    =  30.000000
site_refhgt =   58.0
dxy         =   0.1
undef       = -9999.99
```

UTC offset, if data are provided in local time. For example, if the location is in the 
Western Hemisphere and the time zone is UTC-3, then utc_off should be -3.
**Important**.  If the input data are already in UTC, then set utc_off to zero.
```{r, label='utc-settings'}
utc_off = -3
```

How to interpret the time stamp.  Currently there are three options, as described below.
To help illustrating the cases, we give examples of the beginning and end of average
period of an example time stamp (2021-06-11 08:00) for an hourly average data where the 
high-frequency data were collected at 1-minute resolution.

**imetavg**  | **Description**               | **Beginning**    | **End**
------------ | ----------------------------- | ---------------- | ----------------
1            | End of averaging period       | 2021-06-11 07:01 | 2021-06-11 07:00
2            | Beginning of averaging period | 2021-06-11 08:00 | 2021-06-11 08:59
3            | Middle of averaging period    | 2021-06-11 07:31 | 2021-06-11 08:30

```{r imetavg-setting}
imetavg = 1
```


## Input variables

Here we will provide information about the meteorological driver variables. Our goal is to have
the variables names and units as defined in the table below:

**FATES variable** | **Description**              | **Units**
------------------ | ---------------------------- | --------------------------
**PSRF**           | Atmospheric pressure         | Pa
**TBOT**           | Air temperature              | K
**QBOT**^1^        | Specific humidity            | kg kg^-1^
**RBOT**^1^        | Vapour mixing ratio          | kg kg^-1^
**EBOT**^1^        | Vapour pressure              | Pa
**RH**^1^          | Relative humidity            | fraction (Pa Pa^-1^)
**WIND**           | Wind speed                   | m s^-1^
**FSDS**           | Incident solar irradiance    | W m^-2^
**FLDS**^2^        | Incident longwave irradiance | W m^-2^
**PRECTmms**       | Precipitation rate           | kg m^-2^ s^-1^ (mm s^-1^)

**Notes** 

1. Only one of the humidity variables must be provided. If more than one is provided, the 
script will use only the first one.
2. Incident longwave irradiance (FLDS) is currently not required by ELM or CLM.  
In case you have the data, we suggest to add it to the output as it could be useful for
other purposes.  Otherwise, delete/comment the list entry for FLDS in the following chunk.


First, we define some unit conversion constants for the chunk below
```{r, label='unit-conversion-factors'}
Pa_2_hPa  = 0.01     ; hPa_2_Pa  = 1. / Pa_2_hPa
K_2_degC  = -273.15  ; degC_2_K  = - K_2_degC
kg_2_g    = 1000.    ; g_2_kg    = 1. / kg_2_g
sec_2_hr  = 1./3600. ; hr_2_sec  = 1. / sec_2_hr
sec_2_day = 1./86400.; day_2_sec = 1. / sec_2_day
frac_2_pc = 100.     ; pc_2_frac = 1. / frac_2_pc
```



Use then use the `varinfo` list to standardise the output.  Each list element contains the 
following elements:

* *vfates*.  Variable name from the table above (don't change this value).
* *vlname*.  Long name for variable (don't change this value).
* *vunits*.  Units for the variable (same as the table above, don't change this value).
* *vinput*.  Variable name in the input file.
* *add0*.    Value to be added to the original input 
* *mult*.    Value to be multiplied to the original input

For `add0` and `mult`, except for trivial factors (`add0=0.` or `mult=1`), refrain from typing
"magic numbers" here.  Instead, use the unit conversion parameters described in the chunk above.
In case the sought unit conversion is not available in the previous chunk, feel free to add 
more factors.

```{r, label='var-info-settings'}
n       = 0
varinfo = list()
n       = n + 1
varinfo[[n]] = list( vfates = "PSRF"
                   , vlname = "surface pressure at the tower"
                   , vunits = "Pa"
                   , vinput = "atm.prss"
                   , add0   = 0.
                   , mult   = hPa_2_Pa
                   )#end list
n       = n + 1
varinfo[[n]] = list( vfates = "TBOT"
                   , vlname = "temperature at the tower"
                   , vunits = "K"
                   , vinput = "atm.tmp"
                   , add0   = degC_2_K
                   , mult   = 1.
                   )#end list
n       = n + 1
varinfo[[n]] = list( vfates = "RH"
                   , vlname = "relative humidity at the tower"
                   , vunits = "Pa/Pa"
                   , vinput = "atm.rhv"
                   , add0   = 0.
                   , mult   = pc_2_frac
                   )#end list
n       = n + 1
varinfo[[n]] = list( vfates = "WIND"
                   , vlname = "wind at the tower"
                   , vunits = "m/s"
                   , vinput = "atm.vels"
                   , add0   = 0.
                   , mult   = 1.
                   )#end list
n       = n + 1
varinfo[[n]] = list( vfates = "FSDS"
                   , vlname = "incident solar radiation at the tower"
                   , vunits = "W/m2"
                   , vinput = "rshort.in"
                   , add0   = 0.
                   , mult   = 1.
                   )#end list
n       = n + 1
varinfo[[n]] = list( vfates = "PRECTmms"
                   , vlname = "precipitation rate at the tower"
                   , vunits = "mm/s"
                   , vinput = "rain"
                   , add0   = 0.
                   , mult   = sec_2_hr
                   )#end list
# Longwave radiation. Comment the lines below in case the variable is not available.
n       = n + 1
varinfo[[n]] = list( vfates = "FLDS"
                   , vlname = "incident long wave radiation at the tower"
                   , vunits = "W/m2"
                   , vinput = "rlong.in"
                   , add0   = 0.
                   , mult   = 1.
                   )#end list

# Convert varinfo to a "tibble" object
varinfo  = do.call(what=rbind,args=lapply(X=varinfo,FUN=as_tibble,stringsAsFactors=FALSE))
```


Last, we check that at least one of the humidity variables is listed. In case multiple variables
were provided, they will be all written to the netcdf. Specific humidity will be always written.
If not provided, then we will calculate it and add it to the output.
```{r, label='humid-check'}
humid.vars = c("QBOT","RBOT","EBOT","RH")
is.humid   = any(humid.vars %in% varinfo$vfates)
if (! is.humid){
   cat(" At least one humidity variable must be provided!\n")
   cat(" - QBOT (specific humidity).         Provided = ","QBOT" %in% varinfo$vfates,"\n")
   cat(" - RBOT (water vapour mixing ratio). Provided = ","RBOT" %in% varinfo$vfates,"\n")
   cat(" - EBOT (water vapour pressure).     Provided = ","EBOT" %in% varinfo$vfates,"\n")
   cat(" - RH   (relative humidity).         Provided = ","RH"   %in% varinfo$vfates,"\n")
   stop(" Check that at least one of the humidity variables were provided.")
}#end if
```


This concludes the initial settings.  From this point on, you may not need to change anything,
unless you are debugging or adding new features.

# Data processing


## Met driver loading

Read in the csv file.
```{r, label='csv-load'}
metd_orig = read.csv(file=input_file,header=TRUE,comment.char="",stringsAsFactors=FALSE)
metd_orig = tibble(metd_orig)
metd_orig = mutate( .data  = metd_orig
                  , tstamp = make_datetime( year   = year
                                          , month  = month
                                          , day    = day
                                          , hour   = hour
                                          , min    = min
                                          , tz     = "UTC"
                                          )#end make_datetime
                  )#end mutate
```

## Time standardisation

Find the time difference, and assess the first and last year with full data.
```{r, label='time-info'}
#---- Time interval between observations.
dtstamp = mean(diff(metd_orig$tstamp))

#---- Number of observations per day.
nperday = time_length(make_difftime(day=1),unit="hour") / time_length(x=dtstamp,unit="hour")

#---- Find which years have complete data
datsumm = group_by (.data=metd_orig,year,month) %>% 
          summarise( count  = n()) %>%
          mutate   (expect = days_in_month(make_datetime(year=year,month=month)) * nperday) %>%
          group_by (year) %>%
          summarise(fine = all(count == expect))

#---- Find first and last complete years
allfine = all(datsumm$fine)
yeara   = min(datsumm$year[datsumm$fine])
yearz   = max(datsumm$year[datsumm$fine])

#---- Find and last expected time
tstampa = make_datetime( year=yeara  ,month=1,day=1)
tstampz = make_datetime( year=yearz+1,month=1,day=1)
```


Correct data to be in UTC whilst accounting for the time stamp.  For FATES, the time stamp 
should always correspond to the middle of the averaging window.
```{r, label='local-time-to-UTC'}
if (imetavg == 1){
  dt.utc = make_difftime(hour=utc_off) - 0.5 * dtstamp
}else if (imetavg == 2){
  dt.utc = make_difftime(hour=utc_off) + 0.5 * dtstamp
}else if (imetavg == 3){
  dt.utc = make_difftime(hour=utc_off)
}#end if (imetavg)

metdriver = metd_orig %>% 
            mutate(tstamp = tstamp + dt.utc) %>%
            select(!c(year,month,day,hour,min))
```




With the time zone shift, we may have a few data points off at the beginning or at the end.  
As a simple solution, we just shift these hours to the beginning or the end of the simulation,
just so every year/month has all the data.
```{r, label='time-correct'}
metdriver = metdriver %>% 
            mutate( tstamp  = case_when( tstamp < tstampa ~ tstampz + difftime(tstamp,tstampa)
                                       , tstamp > tstampz ~ tstampa + difftime(tstamp,tstampz)
                                       , TRUE             ~ tstamp
                                       )#end when
                  ) %>% arrange(tstamp)
```


## Unit standardisation for output


Simplify output so variables are in the correct units, and only variables needed for the output 
remain in the output.
```{r, label='simplify-driver'}
met_fates = metdriver %>% 
            select(c(tstamp,varinfo$vinput)) %>%
            rename_at(vars(c("tstamp",varinfo$vinput)), ~ c("time",varinfo$vfates))
            
```

Fix units for all variables:
```{r, label='units-standard'}
# Create a simple addition tibble
add0 = varinfo %>% mutate(id=1) %>% select(c(id,vfates,add0)) %>% 
       spread(vfates,add0) %>% select(-id) %>% relocate(varinfo$vfates)             

# Create a simple multiplication tibble
mult = varinfo %>% mutate(id=1) %>% select(c(id,vfates,mult)) %>% 
       spread(vfates,mult) %>% select (! id) %>% relocate(varinfo$vfates)          

# Fix units. We temporarily remove time from the tibble, so we can add and multiply.
tstamp    = met_fates$time
met_fates = met_fates %>% select(! time)
met_fates = as_tibble(map2(met_fates,add0,function(x,y) if (is.numeric(x)){x+y}else{x}))
met_fates = as_tibble(map2(met_fates,mult,function(x,y) if (is.numeric(x)){x*y}else{x}))
met_fates = met_fates %>% mutate( time = tstamp) %>% relocate(c("time",varinfo$vfates))
```


# Derive specific humidity

This step is only needed if specific humidity was provided. In case not, we derive specific humidity ($q_{v}$) [ $\mathrm{kg_{v}\,kg_{a}^{-1}}$] using one the following relationships (in this order of preference), using the fact that specific humidity is the ratio of mass (or partial density) of water vapour $\rho_{v}$ [$\mathrm{kg_{v}\,m^{-3}}$] and total mass (or density) of air ($\rho_{a}$)  [$\mathrm{kg_{a}\,m^{-3}}$].

* _Mixing ratio_ $r_v$ [ $\mathrm{kg_{v}\,kg_{d}^{-1}}$] is defined as the ratio between partial density of water vapour $\rho_{v}$ and partial density of dry air ($\rho_{d}$) [$\mathrm{kg_{d}\,m^{-3}}$]. Assuming that $\rho_{a} = \rho_{v} + \rho_{d}$, we can relate mixing ratio and specific humidity as: $$\begin{align} & q_{v} = \frac{r_{v}}{r_{v}+1} \end{align}.$$

* To convert from _water vapour pressure_ $e_{v}$ [$\mathrm{Pa}$] to $q_v$, we use the ideal gas equation of a mixture of dry air and water vapour , and account for the difference in molar masses of water vapour ($\mathcal{M}_v = 1.802 \times 10^{-2}\,\mathrm{kg\,mol^{-1}}$)
and dry air ($\mathcal{M}_d = 2.897 \times 10^{-2}\,\mathrm{kg\,mol^{-1}}$): $$\begin{align}
& q_{v} = \frac{\varepsilon{}\,e_{v}}{p - \left(1 - \varepsilon{}\right)\,e_{v}}
\end{align},$$ where $p$ [$\mathrm{Pa}$] is the total air pressure and $\varepsilon = \mathcal{M}_{v} / \mathcal{M}_{d}$.

* _Relative humidity_ ($f_{v}$) [fraction, or $\mathrm{Pa\,Pa^{-1}$] is defined as $f_{v}=e_{v}/e_{v}^{\star}$, where $e_{v}^{\star}$ is the vapour pressure at saturation, which depends on temperature $T$ [$\mathrm{K}$]. To derive specific humidity, we use [@bolton1980mwr] empirical equation for $e_{v}^{\star}$ to derive $e_{v}$: $$\begin{align}
& e_{v} = f_{v}\,e_{0}\,\exp{\left[ \frac{17.67\,\left( T - T_{0}\right)}{T - 29.65} \right]}
\end{align},$$ where $e_{0} = 611.66\,\mathrm{Pa}$ is the vapour pressure of saturation at $T_{0}=273.15\,\mathrm{K}$. We then convert $e_{v}$ to $q_{v}$ using the same equation described above.

```{r, label='define-specific-humidity'}

#---~---
# Define some thermodynamic constants.
#---~---
# Molar mass of water vapour [kg/mol]
mm_h2o  = 0.01801505
# Molar mass of dry air [kg/mol]
mm_dry  = 0.02897
# Molar mass ratio
eps_mol = mm_h2o / mm_dry
# Saturation vapour pressure at 273.15K [Pa]
esat_0C = 611.65685464


if ("QBOT" %in% names(met_fates)){
   # Specific humidity already in the meteorological driver, do nothing.
   cat(" - Specific humidity included in the input data.")
}else if ("RBOT" %in% names(met_fates)){
   # Find specific humidity from mixing ratio
   met_fates = met_fates %>%
               mutate( QBOT = RBOT / (1. + RBOT))
}else if ("EBOT" %in% names(met_fates)){
   # Find specific humidity from vapour pressure
   met_fates = met_fates %>%
               mutate( QBOT = eps_mol * EBOT / (PSRF - (1. - eps_mol) * EBOT))
}else{
   # Find specific humidity from vapour pressure
   met_fates = met_fates %>%
               mutate( ESAT = esat_0C * exp( 17.67 * (TBOT - degC_2_K) / (TBOT - 29.65))
                     , EBOT = RH * ESAT
                     , QBOT = eps_mol * EBOT / (PSRF - (1. - eps_mol) * EBOT)
                     ) %>% select(! c(ESAT,EBOT))
}

# Update or insert specific humidity to the list of output variables.
varinfo = varinfo %>%
          rows_upsert( tibble( vfates = "QBOT"
                             , vlname = "specific humidity at the tower"
                             , vunits = "kg/kg"
                             )#end tibble
                     )#end rows_upsert

```



# Data output


Create paths for output, using FATES convention.
```{r, label='output-path'}
# Labels for describing the data set.
ymd_now = today(tzone = "UTC")
ymd_lab = sprintf("%4.4i%2.2i%2.2i",year(ymd_now),month(ymd_now),day(ymd_now))
dxy_lab = sprintf("%.1fx%.1f",dxy,dxy)

# Tag specific for this site, grid, and version (used for file and path names).
site_tag  = paste0(xid,"_",dxy_lab,"_v",dat_version,"_c",ymd_lab)


# Path to where we will write the drivers.
metd_path = file.path(here_path,site_tag,"CLM1PT_data")

# Create directory
dummy = dir.create(path=metd_path  ,showWarnings=FALSE,recursive=TRUE)
```

Group data by month and year

```{r, label='prepare-output'}
# Split data by month/year
met_output = met_fates %>% 
             mutate( year = year(time), month=month(time)) %>%
             group_split(year,month,.keep=FALSE)

# Count groups
nmet = length(met_output)

```

Generate edge information for the single data point.
```{r, label='geo-find'}
outlon  = site_lon %% 360.
outlat  = site_lat
edge_w  = (site_lon - dxy) %% 360.
edge_e  = (site_lon + dxy) %% 360.
edge_s  = site_lat - dxy
edge_n  = site_lat + dxy
```

Define the template for the global attributes.
```{r, label='global-attributes', message=FALSE, results='hide'}

# Define the code developer information (indirect way so the email is not visible).
developer_name  = c( 111L, 103L, 110L, 111L,  76L,  32L, 115L, 111L,  99L, 114L,  97L,  77L)
developer_email = c( 118L, 111L, 103L,  46L, 108L,  98L, 108L,  64L, 111L, 103L, 110L, 111L
                   , 108L, 109L)


# Define the template.  We will update the title in each time step.
att_template = list( title          = "To be replaced when looping through months"
                   , version        = dat_version
                   , date_created   = paste0(as.character(now(tzone="UTC")), "UTC")
                   , source_code    = "make_fates_met_driver.Rmd"
                   , code_notes     = "Meteorological drivers compatible with ELM-FATES and CLM-FATES"
                   , code_developer = paste0( intToUtf8(rev(developer_name))
                                            ," <"
                                            , intToUtf8(rev(developer_email))
                                            ,">"
                                            )#end paste0
                   , file_author    = paste0(author_name," <",author_email,">")
                   , data_provider  = paste0(datprov_name," <",datprov_email,">")
                   , usage_notes    = data_usage_notes
                   )#end list
```


Loop through every group, and generate the output file.
```{r, label='write-output', message=FALSE, results='hide'}
for (m in sequence(nmet)){
   # Copy the subset to a local variable.
   met_this = met_output[[m]]
   nthis    = nrow(met_this)

   # Find first time for this month
   year_this   = unique(year (met_this$time))
   month_this  = unique(month(met_this$time))
   first_this  = make_date(year_this,month_this,1)

   
   # Extract time, and turn it into a difference in days
   tsince = as.numeric(difftime(met_this$time,first_this,units="days"))

      
   # Create label for this month
   year_this  = unique(year (met_this$time))
   month_this = unique(month(met_this$time))
   when_lab   = sprintf("%4.4i-%2.2i",year_this,month_this)

   
   # File name
   nc_base = paste0(when_lab,".nc")
   nc_file = file.path(metd_path,nc_base)
   cat(" + Write output for ",when_lab," (",nc_base,").\n",sep="")

   # In case file exists, it will be re-created.
   if (file.exists(nc_file)) file.remove(nc_file)

   # Add dimensions: longitude, latitude, and time. We do not automatically create the 
   # dimension variable for time because R would create it in double precision.  Instead,
   # we append variable time manually.
   xx  = ncdim_def( name="lon"   ,units="",vals=1L               ,create_dimvar=FALSE)
   yy  = ncdim_def( name="lat"   ,units="",vals=1L               ,create_dimvar=FALSE)
   tt  = ncdim_def( name="time"  ,units="",vals=seq_along(tsince),create_dimvar=FALSE)
   ss  = ncdim_def( name="scalar",units="",vals=1L               ,create_dimvar=FALSE)

   # List of dimensions, useful for setting variables.   
   nc_xy  = list   (xx,yy)
   nc_xyt = list(xx,yy,tt)
   nc_t   = list      (tt)
   nc_s   = list(ss)
   xy     = c(1,1)
   xyt    = c(1,1,nthis)

   # Start list with variables. First we put the coordinates
   nc_vlist        = list()
   nc_vlist$LONGXY = ncvar_def( name     = "LONGXY"
                              , units    = "degrees_east"
                              , dim      = nc_xy
                              , missval  = undef
                              , longname = "longitude"
                              )#end ncvar_def
   nc_vlist$LATIXY = ncvar_def( name     = "LATIXY"
                              , units    = "degrees_north"
                              , dim      = nc_xy
                              , missval  = undef
                              , longname = "latitude"
                              )#end ncvar_def
   nc_vlist$ZBOT   = ncvar_def( name     = "ZBOT"
                              , units    = "m"
                              , dim      = nc_xy
                              , missval  = undef
                              , longname = "observational height"
                              )#end ncvar_def
   nc_vlist$EDGEW  = ncvar_def( name     = "EDGEW"
                              , units    = "degrees_east"
                              , dim      = nc_s
                              , missval  = undef
                              , longname = "western edge in atmospheric data"
                              )#end ncvar_def
   nc_vlist$EDGEE  = ncvar_def( name     = "EDGEE"
                              , units    = "degrees_east"
                              , dim      = nc_s
                              , missval  = undef
                              , longname = "eastern edge in atmospheric data"
                              )#end ncvar_def
   nc_vlist$EDGES  = ncvar_def( name     = "EDGES"
                              , units    = "degrees_north"
                              , dim      = nc_s
                              , missval  = undef
                              , longname = "southern edge in atmospheric data"
                              )#end ncvar_def
   nc_vlist$EDGEN  = ncvar_def( name     = "EDGEN"
                              , units    = "degrees_north"
                              , dim      = nc_s
                              , missval  = undef
                              , longname = "northern edge in atmospheric data"
                              )#end ncvar_def
   nc_vlist$time   = ncvar_def( name     = "time"
                              , units    = paste0( "days since ",as.character(first_this)
                                                 , " 00:00:00"
                                                 )#end paste0
                              , dim      = nc_t
                              , missval  = undef
                              , longname = "observation time"
                              )#end ncvar_def

   # Loop through FATES met drivers, add them   
   for (v in seq_along(varinfo[[1]])){
      # Handy shorter names
      v_vfates = varinfo$vfates[v]
      v_vlname = varinfo$vlname[v]
      v_vunits = varinfo$vunits[v]

      #Add variable information
      nc_vlist[[v_vfates]] = ncvar_def( name     = v_vfates
                                      , units    = v_vunits
                                      , dim      = nc_xyt
                                      , missval  = undef
                                      , longname = v_vlname
                                      )#end ncvar_def
   }#end for (v in seq_along(varinfo[[1]]))

   # Create file
   nc_conn = nc_create(filename=nc_file,vars=nc_vlist,verbose=FALSE)

   #---~---
   # Put coordinates, tower height and attributes to the netcdf
   #---~---
   # Longitude, append time-invariant tag
   dummy = ncvar_put(nc=nc_conn,varid="LONGXY",vals=array(data=outlon     ,dim=xy))
   dummy = ncatt_put(nc=nc_conn,varid="LONGXY",attname="mode"    ,attval="time-invariant")
   # Latitude, append time-invariant tag
   dummy = ncvar_put(nc=nc_conn,varid="LATIXY",vals=array(data=outlat     ,dim=xy))
   dummy = ncatt_put(nc=nc_conn,varid="LATIXY",attname="mode"    ,attval="time-invariant")
   # Tower elevation above ground level, append time-invariant tag
   dummy = ncvar_put(nc=nc_conn,varid="ZBOT"  ,vals=array(data=site_refhgt,dim=xy))
   dummy = ncatt_put(nc=nc_conn,varid="ZBOT"  ,attname="mode"    ,attval="time-invariant")
   # Western edge, append time-invariant tag
   dummy = ncvar_put(nc=nc_conn,varid="EDGEW" ,vals=edge_w)
   dummy = ncatt_put(nc=nc_conn,varid="EDGEW" ,attname="mode"    ,attval="time-invariant")
   # Eastern edge, append time-invariant tag
   dummy = ncvar_put(nc=nc_conn,varid="EDGEE" ,vals=edge_e)
   dummy = ncatt_put(nc=nc_conn,varid="EDGEE" ,attname="mode"    ,attval="time-invariant")
   # Southern edge, append time-invariant tag
   dummy = ncvar_put(nc=nc_conn,varid="EDGES" ,vals=edge_s)
   dummy = ncatt_put(nc=nc_conn,varid="EDGES" ,attname="mode"    ,attval="time-invariant")
   # Northern edge, append time-invariant tag
   dummy = ncvar_put(nc=nc_conn,varid="EDGEN" ,vals=edge_n)
   dummy = ncatt_put(nc=nc_conn,varid="EDGEN" ,attname="mode"    ,attval="time-invariant")
   # Time, append calendar type.
   dummy = ncvar_put(nc=nc_conn,varid="time"  ,vals=tsince)
   dummy = ncatt_put(nc=nc_conn,varid="time"  ,attname="calendar",attval="gregorian")
   #---~---

   # Put variables to the netcdf
   for (v in seq_along(varinfo[[1]])){
      # Handy shorter names
      v_vfates = varinfo$vfates[v]
      v_vlname = varinfo$vlname[v]
      v_vunits = varinfo$vunits[v]

      #Add variable information
      dummy = ncvar_put( nc    = nc_conn
                       , varid = v_vfates
                       , vals  = array(data=met_fates[[v_vfates]],dim=xyt)
                       )#end ncvar_put

      #Add attribute to highlight this is time-dependent
      dummy = ncatt_put( nc      = nc_conn
                       , varid   = v_vfates
                       , attname = "mode"
                       , attval  = "time-dependent")
   }#end for (v in seq_along(varinfo[[1]]))

   # Add title specific for this month/year.
   nc_title   = paste0( "Meteorological forcing for ",site_desc
                      , "(",month.abb[month_this]," ",year_this,")"
                      )#end paste0
   att_global = modifyList( x = att_template, val = list( title = nc_title ))

   
   # Loop through global attributes
   for (l in seq_along(att_global)){
      # Current attribute information
      att_name  = names(att_global)[l]
      att_value = att_global[[l]]

      # Add attribute 
      dummy = ncatt_put(nc=nc_conn,varid=0,attname=att_name,attval=att_value)
   }#end for (l in seq_along(att_global))


   # Close the file
   dummy = nc_close(nc_conn)
}#end for (m in sequence(nmet))
```
# References

