---
title: "Processor for MODIS LAI time series at points of interest"
author: "Marcos Longo"
date: "19-Oct-2021"
output: html_document
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

This document shows how to generate time series of MODIS-derived leaf area index and FPAR for points of interest. The code will aggregate pixels near the point of interest, restrict the vegetation to sought classes only, and process only the data with the highest quality used. 

This script relies on data downloaded through [AppEEARS](https://lpdaacsvc.cr.usgs.gov/appeears/).  There you will need to request the following data centred around the points of interest:

**Product**   | **Layer**           | **Description**                                 | **Spatial Resolution** | **Temporal Resolution**
------------- | ------------------- | ----------------------------------------------- | ---------------------- | ------------------------
**MCD12Q1**   | **LC_Type1**        | Annual IGBP land cover classificat0ion	        | 500 m                  | 1 year
**MCD12Q1**   | **QC**              | Quality flags for land cover classificat0ion     | 500 m                  | 1 year
**MCD15A2H**  | **Lai_500m**        | Leaf area index                                 | 500 m                  | 8 days
**MCD15A2H**  | **LaiStdDev_500m**  | Standard deviation of leaf area index           | 500 m                  | 8 days
**MCD15A2H**  | **Fpar_500m**       | Fraction of Photosynthetically Active Radiation | 500 m                  | 8 days
**MCD15A2H**  | **FparStdDev_500m** | Standard deviation of FPAR                      | 500 m                  | 8 days
**MCD15A2H**  | **FparLai_QC**      | Quality control flags for FPAR and LAI          | 500 m                  | 8 days
**MCD15A2H**  | **FparExtra_QC**    | Extra detail quality for FPAR and LAI           | 500 m                  | 8 days
|

This script assumes that the output data are saved in **NetCDF-4** format, using either **geographic coordinates** (preferred) or _MODIS Sinusoidal_ coordinates. Make sure to select these settings when requesting data from AppEEARS. The script was written for Collection 6. 

# Reset session

Use this chunk to fully reset R.
```{r, label = 'reset-R',message=FALSE, results='hide'}
# Unload all packages except for the R default ones
plist = names(sessionInfo()$otherPkgs)
if (length(plist) > 0){
   dummy = sapply(X=paste0("package:",plist),FUN=detach,character.only=TRUE,unload=TRUE)
}#end if (length(plist) > 0)


# Remove all variables
rm(list=ls())

# Reset warnings
options(warn=0)

# Close all plots
invisible(graphics.off())

# Clean up
invisible(gc())
```


# Main settings

## Path and file locat0ion

Set paths and files for input and output.

* **home_path**. Typically the user's home path.  Useful for building other paths. `path.expand("~")` typically works for all users.
* **main_path**.   The main path for processing data. The output path will be generated here.
* **util_path**.   Directory with useful functions and packages.
* **input_path**.  The locat0ion where existing surface and domain netcdf files are locat0ed
* **output_path**. The main output path for the data. A sub-directory for this site will be created.
* **plot_path**.   The path for plots.

```{r, label='path-settings'}
home_path   = path.expand("~")
util_path   = file.path(home_path,"Util","RUtils")
main_path   = file.path(home_path,"Data","MODIS_LAI")
input_path  = file.path(main_path,"InputData")
output_path = file.path(main_path,"MCD15A2H_Site")
plot_path   = file.path(main_path,"Figures")
```



Set the sites to process images. We put the information in a tibble object `poi_list` containing the following variables.

* **site**. Site name (it must match the name of the sub-directory of `input_path`.  Normally this is the same name given to the data retrieval from AppEEARS)
* **output**. Output identifier, FATES style. This is useful for adding the data to existing folders.
* **lon**. Longitude of the point of interest.  The LAI output will pool data from multiple pixels centred around this longitude (defined by `sample_radius`).
* **lat**. Latitude of the point of interest.  The LAI output will pool data from multiple pixels centred around this latitude.(defined by `sample_radius`).
* **dxy**. Intended grid size (we will generate a 1x1-point pseudo-grid for FATES).
* **geographic**. Coordinate system used for this site.  If `TRUE`, we assum geographic coordinates (i.e., longitude and latitude).  If `FALSE`, we assume MODIS sinusoidal.
* **igbp_include**. A character with the classes to include in the sample (represented by letters, case insensitive). Use the table below to guide the selection.
* **colour**. A character with the colour to be used for each site.

The following table shows which classes are included when each of the flags are set to true

LC_Type1 Class | Description                           | igbp_include letter
-------------- | ------------------------------------- | --------------------
 1             | Evergreen Needleleaf Forests          | a
 2             | Evergreen Broadleaf Forests           | b
 3             | Deciduous Needleleaf Forests          | c
 4             | Deciduous Broadleaf Forests           | d
 5             | Mixed Forests                         | e
 6             | Closed Scrublands                     | f
 7             | Open Scrublands                       | g
 8             | Woody Savannahs                       | h
 9             | Savannahs                             | i
 10            | Grasslands                            | j
 11            | Permanent Wetlands                    | k
 12            | Croplands                             | l
 13            | Urban and Built-up Lands              | m
 14            | Cropland Natural Vegetation Mosaics   | n
 15            | Permanent Snow and Ice                | o
 16            | Barren                                | p
 17            | Water Bodies                          | q
|

```{r, label='poi-list'}
# List of polygons to process.
poi_list = tidyr::tribble( ~site      , ~output                              , ~lon   ,    ~lat, ~dxy, ~geographic, ~igbp_include,   ~colour
                         , "Paracou"  , "1x1pt-paracouGUF_v1.6_c20210913"    , -52.912,   5.282,  1.0,        TRUE,          "bd", "#8DA0CB"
                         , "Tanguro"  , "1x1pt-tanguroMT-BR_v1.2_c20210913"  , -52.409, -13.081,  1.0,        TRUE,          "bd", "#66C2A5"
                         , "Petrolina", "1x1pt-petrolinaPE-BR_v1.2_c20210913", -40.370,  -9.165,  1.0,        TRUE,      "bdfghi", "#FC8D62"
                         )#end tribble

# Number of polygons to process
n_poi_list = nrow(poi_list)
```

Additional settings for the output file

* **dat_version**. Version of this data set.
* **undef_out**. A number to denote missing values in the output (NetCDF) file. This is needed by NetCDF.

```{r, label='output-settings'}
dat_version = "1.6"
undef_out   = -9999.99
```



Additional settings for data processing:

* **sample_radius**. The maximum radial distance to the reference point that shall be used for the sampling.
* **ci_range**. Range of data to show for time series.
* **alpha_ribbon**. Opaqueness level for ribbon (0 means transparent, 1 means fully opaque).

```{r, label='appeears-settings'}
# Set maximum distance from the point of interest that we will consider for sampling [in kilometres]
sample_radius = 20.

# Which quality flags to consider data high quality
qcfine_mcd15a2h  = c(0,2,32,34) # Fine classes: FparLai_QC (MCD15A2H)
exfine_mcd15a2h = c(0,128)      # Fine classes: FparExtra_QC (MCD15A2H)
qcfine_mcd12q1   = c(0)         # Fine classes: QC (MCD12Q1)

#  Set the range of data to show for time series.
ci_range = 2*pnorm(1)-1

# Opaqueness factor for ribbon
alpha_ribbon = 0.4
```


General plot options for `ggplot`

* **gg_device**. A vector with file types for figures. These should be extensions of common file formats (e.g., pdf, eps, tif, png, jpg).  The complete list of options can be found in [ggsave](https://www.rdocumentation.org/packages/ggplot2/versions/0.9.0/topics/ggsave).
* **gg_depth**. The resolution of this figure (in pixels per inch), in case the figure is saved in raster format (e.g., tif, png, jpg). Ignored for vector format (e.g., pdf, eps).
* **gg_ptsz**. The typical size for fonts (in pt). Larger sizes make it more readable, but shrink the plotting area.
* **gg_width**. The width of the output files.  The units are defined by `gg_units`.
* **gg_height**. The height of the output files.  The units are defined by `gg_units`.
* **gg_units**. Units for `gg_width` and ``gg_height`.  Acceptable units are `"in"` (inches), `"cm"` (centimetres) and `"mm"` (millimetres).
* **gg_screen**. Show images in the documentation as well. If `FALSE`, the image files will be generated, but not shown in the knitted documentation.
* **gg_tmft**. Format for time strings in the time series. Check [scale_x_datetime](https://www.rdocumentation.org/packages/ggplot2/versions/1.0.0/topics/scale_x_datetime) for additional details.
* **gg_ncolours**. Number of colours to use in heat maps (such as soil time series and plots by DBH class and PFT)

```{r, label='set-ggplot',message=FALSE,results='hide'}
gg_device   = c("pdf") # Output devices to use (Check ggsave for acceptable formats)
gg_depth    = 300      # Plot resolution (dpi)
gg_ptsz     = 18       # Font size
gg_width    = 11.0     # Plot width (units below)
gg_height   = 8.5      # Plot height (units below)
gg_units    = "in"     # Units for plot size
gg_screen   = TRUE     # Show plots on screen as well?
gg_tfmt     = "%Y"     # Format for time strings in the time series
gg_ncolours = 129      # Number of node colours for heat maps.
gg_fleg     = 1./6.    # Fraction of plotting area dedicat0ed for legend

# Number of output types.
ndevice = length(gg_device)
```



This concludes the initial settings.  From this point on, you may not need to change anything,
unless you are debugging or adding new features.

# Data processing


## Initial settings

First, we load some useful packages and tools, using the R script `load.everything.r`. This will load all the R scripts at the `util_path` directory.
```{r, label='load-everything'}
source(file.path(util_path,"load.everything.r"),chdir=TRUE)
```

## Define output paths

```{r, label='set-paths'}
# Create output paths
dummy = dir.create(path=output_path,recursive=TRUE,showWarnings=FALSE)
dummy = dir.create(path=plot_path  ,recursive=TRUE,showWarnings=FALSE)
```


## Load data

Here we loop through all the sites and load the data.  First, we make sure all data are in geographic coordinates (even if the data were retrieved in MODIS sinusoidal coordinates).  Then we eliminate points with subpar quality from the analysis.  We then aggregate the 8-day data into monthly, to increase the sample size.  We then append the information to a global tibble, which will be used to make the site-level time series.

We always convert the results to geographic coordinates before aggregating and sub-sampling.

```{r, label='load-modis'}

# Make sure no connection is open before reading the data
if ("nc_conn" %in% ls()) dummy = nc_close(nc=nc_conn)

# Initialise the data set
emean = NULL

# Loop through sites, and load the data.
for (p in sequence(n_poi_list)){
   # Useful short names
   p_site       = poi_list$site      [p]
   p_output     = poi_list$output    [p]
   p_lon        = poi_list$lon       [p]
   p_lat        = poi_list$lat       [p]
   p_geographic = poi_list$geographic[p]
   p_igbp       = match(unlist(strsplit(poi_list$igbp_include[p],split="")),letters)
   p_input_path = file.path(input_path,p_site)
   cat0(" + Retrieve data from ",p_site,".")

   # Find data sets with MCD12Q1 and MCD15A2H data
   p_mcd12q1    = list.files(path=p_input_path,pattern="^MCD12Q1(.+)\\.nc$")
   p_mcd15a2h   = list.files(path=p_input_path,pattern="^MCD15A2H(.+)\\.nc$")
   
   # Make sure both files were unambiguously found
   if ( (length(p_mcd12q1) != 1 ) || (length(p_mcd15a2h) != 1 ) ){
      cat0( " At site ",p_site,".")
      cat0( " Directory ",p_input_path,".")
      cat0( " This directory should contain EXACTLY one file with MCD12Q1 data,")
      cat0( "    and one file with MCD15A2H.  Instead, this is what was found." )
      cat0( " - Number of MCD12Q1 files  = ",length(p_mcd12q1 ),".")
      cat0( " - Number of MCD15A2H files = ",length(p_mcd15a2h),".")
      stop(paste0(" Fix the input path for this site before proceeding."))
   }#end if ( (length(p_mcd12q1) != 1 ) || (length(p_mcd15a2h) != 1 ) )

   # Read the MCD15A2H data
   cat0("   - Read data from file ",p_mcd15a2h,".")
   nc_conn  = nc_open(filename=file.path(p_input_path,p_mcd15a2h))

   # Get dimensions. 
   cat0("     > Get dimensions.")
   if ( p_geographic ){
      # Get dimensions
      nc_nx    = nc_conn$dim$lon$len
      nc_ny    = nc_conn$dim$lat$len
      nc_nt    = nc_conn$dim$time$len
      nc_yswap = rev(sequence(nc_ny))
   }else{
      # Get dimensions
      nc_nx    = nc_conn$dim$xdim$len
      nc_ny    = nc_conn$dim$ydim$len
      nc_nt    = nc_conn$dim$time$len
      nc_yswap = rev(sequence(nc_ny))
   }#end if ( p_geographic )

      
   # Retrieve the times. This happens in a few steps. 
   cat0("     > Find times.")
   # Step 1.  Get first time.
   nc_time0 = nc_conn$dim$time$units
   nc_time0 = gsub(pattern="days since ",replacement="" ,x=nc_time0)
   nc_time0 = gsub(pattern=":"          ,replacement=" ",x=nc_time0)
   nc_time0 = gsub(pattern="-"          ,replacement=" ",x=nc_time0)
   nc_time0 = as.numeric(unlist(strsplit(x=nc_time0,split=" ")))
   nc_time0 = make_date(year=nc_time0[1],month=nc_time0[2],day=nc_time0[3])
   # Step 3. Get the times. Note that MODIS provides data in Julian calendar, not Gregorian. 
   #         Because 2000 was a leap year in Gregorian and Julian, this distinction does not
   #         matter, but could matter in 2100 or if adapting this script to other data sets.
   nc_times = nc_time0 + days(nc_conn$dim$time$vals)
   
   # Retrieve coordinates depending on the reference system
   if ( p_geographic){
      cat0("     > Retrieve longitude/latitude coordinates.")
      nc_lon = nc_conn$dim$lon$vals
      nc_lat = rev(nc_conn$dim$lat$vals)

      # Initialise the tibble
      p_datum   = tibble( ixy    = rep(x=sequence(nc_nx*nc_ny)    ,times=nc_nt)
                        , lon    = rep(x=rep(x=nc_lon,times=nc_ny),times=nc_nt)
                        , lat    = rep(x=rep(x=nc_lat,each =nc_nx),times=nc_nt)
                        , time   = rep(x=nc_times,each=nc_nx*nc_ny)
                        , year   = year(time)
                        , month  = month(time)
                        , day    = day(time)
                        )#end tibble

      # Remove lubridate time
      p_datum   = p_datum %>% select(c(ixy,lon,lat,year,month,day))
      
   }else{
      cat0("     > Retrieve sinusoidal coordinates.")
      # Get sinusoidal coordinates
      nc_xsinus = nc_conn$dim$xdim$vals
      nc_ysinus = rev(nc_conn$dim$ydim$vals)
      
      # Get the radius
      nc_erad   = ncatt_get(nc=nc_conn,varid="crs",attname="radius_of_sphere")$value

      # Initialise the tibble
      p_datum   = tibble( ixy    = rep(x=sequence(nc_nx*nc_ny)       ,times=nc_nt)
                        , xsinus = rep(x=rep(x=nc_xsinus,times=nc_ny),times=nc_nt)
                        , ysinus = rep(x=rep(x=nc_ysinus,each =nc_nx),times=nc_nt)
                        , time   = rep(x=nc_times,each=nc_nx*nc_ny)
                        , year   = year(time)
                        , month  = month(time)
                        , day    = day(time)
                        , lat    = 180.*ysinus/nc_erad/pi
                        , lon    = 180.*xsinus/nc_erad/pi/cos(lat*pi/180.)
                        )#end tibble

      # Remove sinusoidal information and lubridate time.
      p_datum   = p_datum %>% select(c(ixy,lon,lat,year,month,day))
   }#end if ( p_geographic)


   # Calculate the distance to central point.
   cat0("     > Find distances to reference point. ")
   p_datum = p_datum %>%
             mutate( dist = c( rdist.earth( x1    = cbind(p_datum$lon,p_datum$lat)
                                          , x2    = cbind(p_lon,p_lat)
                                          , miles = FALSE
                                          )#end rdist.earth
                             )#end c
                   )#end mutate

   # Retrieve the LAI, FPAR, and QC fields   
   cat0("     > Load MCD15A2H variables.")
   p_datum = p_datum %>% 
             mutate( lai     = c(ncvar_get(nc=nc_conn,varid="Lai_500m"       )[,nc_yswap,])
                   , lai_sd  = c(ncvar_get(nc=nc_conn,varid="LaiStdDev_500m" )[,nc_yswap,])
                   , fpar    = c(ncvar_get(nc=nc_conn,varid="Fpar_500m"      )[,nc_yswap,])
                   , fpar_sd = c(ncvar_get(nc=nc_conn,varid="FparStdDev_500m")[,nc_yswap,])
                   , qcflag  = c(ncvar_get(nc=nc_conn,varid="FparLai_QC"     )[,nc_yswap,])
                   , qcextra = c(ncvar_get(nc=nc_conn,varid="FparExtra_QC"   )[,nc_yswap,])
                   )#end mutate
   
   # Close the connection.
   cat0("     > Close connection.")
   dummy = nc_close(nc = nc_conn)

   # Read the MCD12Q1 data
   cat0("   - Read data from file ",p_mcd12q1,".")
   nc_conn  = nc_open(filename=file.path(p_input_path,p_mcd12q1))

      if ( p_geographic ){
      # Get dimensions
      nc_nx12    = nc_conn$dim$lon$len
      nc_ny12    = nc_conn$dim$lat$len
      nc_nt12    = nc_conn$dim$time$len
   }else{
      # Get dimensions
      nc_nx12    = nc_conn$dim$xdim$len
      nc_ny12    = nc_conn$dim$ydim$len
      nc_nt12    = nc_conn$dim$time$len
   }#end if ( p_geographic )


   # Check dimensions. Both the MCD15A2H and MCD12Q1 data sets should have the same spatial dimensions. 
   cat0("     > C dimensions.")
   
   if ( (nc_nx12 != nc_nx) || (nc_ny12 != nc_ny) ){
      cat0("-------------------------------------------------------")
      cat0(" MCD12Q1 and MCD15A2H files are spatially incompatible!")
      cat0("-------------------------------------------------------")
      cat0(" Input file path           = ",p_input_path             )
      cat0(" MCD12Q1 dimensions  (x;y) = (",nc_nx12,";",nc_ny12,")" )
      cat0(" MCD15A2H dimensions (x;y) = (",nc_nx  ,";",nc_ny  ,")" )
      cat0("-------------------------------------------------------")
      stop(" MCD12Q1 and MCD15A2H files should come from the same AppEEARS order!")
   }#end if ( (nc_conn$dim$xdim$len != nc_nx) || (nc_conn$dim$xdim$len != nc_ny) )


   # Load MCD12Q1 variables.
   cat0("     > Load MCD12Q1 variables from most recent year.")
   p_datum = p_datum %>%
             mutate( igbp   = rep(c(ncvar_get(nc=nc_conn,varid="LC_Type1")[,nc_yswap,nc_nt12]),each=nc_nt)
                   , qc12   = rep(c(ncvar_get(nc=nc_conn,varid="QC"      )[,nc_yswap,nc_nt12]),each=nc_nt)
                   )#end mutate

   # Eliminate data that had lower quality.
   cat0("     > Remove data with lower quality.")
   p_datum = p_datum                                %>%
             filter( qcflag  %in% qcfine_mcd15a2h ) %>%
             filter( qcextra %in% exfine_mcd15a2h ) %>%
             filter( qc12    %in% qcfine_mcd12q1  )

   #    Find the mediabs by month and year for each pixel. We use medians instead of averages 
   # because poor-quality data seem to be mostly biased low, so using medians should mitigate 
   # this effect.
   cat0("     > Find pixel-based medians by month and year.")
   p_emean = p_datum                                                 %>%
             group_by(year,month,ixy)                                %>%
             summarise( lon     = mean(lon ,na.rm=TRUE)
                      , lat     = mean(lat ,na.rm=TRUE)
                      , dist    = mean(dist,na.rm=TRUE)
                      , lai     = median(lai ,na.rm=TRUE)
                      , lai_sd  = sqrt(median(lai_sd^2 ,na.rm=TRUE))
                      , fpar    = median(fpar,na.rm=TRUE)
                      , fpar_sd = sqrt(median(fpar_sd^2,na.rm=TRUE))
                      , igbp    = median(igbp,na.rm=TRUE)          ) %>%
             ungroup()                                               %>%
             mutate   ( lc_use  = igbp %in% p_igbp )                 %>%
             mutate   ( site  = p_site
                      , lon0  = p_lon
                      , lat0  = p_lat )                              %>%
             select(site,lon0,lat0,ixy,year,month,lon,lat,dist
                   ,lai,lai_sd,fpar,fpar_sd,igbp,lc_use      )       %>%
             arrange(ixy,year,month)
   
   
      
   # Close the connection.
   cat0("     > Close connection.")
   dummy = nc_close(nc = nc_conn)

   
   # Append data to the global data 
   emean = rbind( emean, p_emean)
}#end for (p in sequence(n_poi_list))
```

## Spatial average

Here all the data are aggregated by month, but not spatially. In this step, we keep only the pixels that are within the pre-defined maximum distance from the point of interest and that has one of the classes that can be aggregated, and apply the spatial average for each month and year. 

```{r, label='modis-spatial-aggr'}
# Find the lower and upper quantiles
quant_lwr = 0.5 - 0.5 * ci_range
quant_upr = 0.5 + 0.5 * ci_range


# Filter data and find the time series for each site.
cat0(" + Find time series by site.")
poi_emean = emean                                                                               %>%
            filter ( ( dist <= sample_radius ) & lc_use )                                       %>%
            mutate ( lai_wgt  = ifelse(test=lai_sd  %gt% 0., yes = 1./lai_sd^2 , no = 0.)
                   , fpar_wgt = ifelse(test=fpar_sd %gt% 0., yes = 1./fpar_sd^2, no = 0.) )     %>%
            group_by (site,year,month)                                                          %>%
            summarise( lai_qmid   = median(x=lai,na.rm=TRUE)
                     , lai_sdev   = sqrt( median(x=lai_sd^2,na.rm=TRUE) )
                     , lai_n      = sum( is.finite(lai) & is.finite(lai_sd) )
                     , fpar_qmid  = median(x=fpar,na.rm=TRUE)
                     , fpar_sdev  = sqrt(median(x=fpar_sd^2,na.rm=TRUE))
                     , fpar_n     = sum( is.finite(fpar) & is.finite(fpar_sd) )           )     %>%
            ungroup()                                                                           %>%
            mutate   ( lai_qlwr   = pmax(0., qnorm(p=quant_lwr,mean=lai_qmid ,sd=lai_sdev) )
                     , lai_qupr   = qnorm(p=quant_upr,mean=lai_qmid ,sd=lai_sdev)
                     , fpar_qlwr  = pmax(0., qnorm(p=quant_lwr,mean=fpar_qmid,sd=fpar_sdev) )
                     , fpar_qupr  = pmin(1., qnorm(p=quant_upr,mean=fpar_qmid,sd=fpar_sdev) ) ) %>%
            mutate ( when = make_date(year=year,month=month))                                   %>%
            select_at(all_of( c("site","when"
                               ,"lai_qlwr","lai_qmid","lai_qupr","lai_sdev","lai_n"
                               ,"fpar_qlwr","fpar_qmid","fpar_qupr","fpar_sdev","fpar_n")))   %>%
            arrange(site,when)

cat0(" + Find mean annual cycle by site.")
poi_mmean = poi_emean                                                                           %>%
            mutate( month    = month(when)
                  , lai_wgt  = ifelse(test=lai_sdev  %gt% 0., yes = 1./lai_sdev^2 , no = 0.)
                  , fpar_wgt = ifelse(test=fpar_sdev %gt% 0., yes = 1./fpar_sdev^2, no = 0.) )  %>%
            group_by (site,month)                                                               %>%
            summarise( lai_qmid   = mean(x=lai_qmid,na.rm=TRUE)
                     , lai_sdev   = sqrt( mean(x=lai_sdev^2,na.rm=TRUE) )
                     , lai_n      = sum( is.finite(lai_qmid) & is.finite(lai_sdev) )
                     , fpar_qmid  = mean(x=fpar_qmid,na.rm=TRUE)
                     , fpar_sdev  = sqrt(mean(x=fpar_sdev^2,na.rm=TRUE))
                     , fpar_n     = sum( is.finite(fpar_qmid) & is.finite(fpar_sdev) )     )    %>%
            ungroup()                                                                           %>%
            mutate   ( lai_qlwr   = pmax(0., qnorm(p=quant_lwr,mean=lai_qmid ,sd=lai_sdev) )
                     , lai_qupr   = qnorm(p=quant_upr,mean=lai_qmid ,sd=lai_sdev)
                     , fpar_qlwr  = pmax(0., qnorm(p=quant_lwr,mean=fpar_qmid,sd=fpar_sdev) )
                     , fpar_qupr  = pmin(1., qnorm(p=quant_upr,mean=fpar_qmid,sd=fpar_sdev) ) ) %>%
            arrange(site,month)
```


# Output plots

First, we plot the time series of LAI for the three sites.

```{r, label='plot-emean'}
cat0(" + Plot the time series of LAI for all sites by month and year.")

# Retrieve useful elements from the POI list
p_site   = poi_list$site
p_colour = poi_list$colour

# First, turn sites into factors
show_emean = poi_emean %>% mutate( site = factor(site,levels=poi_list$site))

# Plot time series with the range
gg_emean = ggplot( data    = show_emean
                 , mapping = aes_string( x      = "when"
                                       , y      = "lai_qmid"
                                       , ymin   = "lai_qlwr"
                                       , ymax   = "lai_qupr"
                                       , group  = "site"
                                       , colour = "site"
                                       , fill   = "site"
                                       )#end aes_string
                 )#end ggplot
gg_emean = gg_emean + scale_colour_manual(name="",aesthetics="colour",labels=p_site,values=p_colour)
gg_emean = gg_emean + scale_colour_manual(name="",aesthetics="fill"  ,labels=p_site,values=p_colour)

# Add ribbons with lower and upper range, and also the medians as lines
gg_emean = gg_emean + geom_ribbon( alpha = alpha_ribbon, colour = "transparent", show.legend=TRUE)
gg_emean = gg_emean + geom_line( lwd = 1.0, show.legend=TRUE)

# Add local annotation
gg_emean = gg_emean + labs(title=element_blank())
gg_emean = gg_emean + scale_x_date(date_labels=gg_tfmt)
gg_emean = gg_emean + xlab(element_blank())
gg_emean = gg_emean + ylab( desc.unit( desc = "MODIS Leaf Area Index (MCD15A2H)"
                                     , unit = untab$m2lom2
                                     , twolines = TRUE
                                     )#end desc.unit
                          )#end ylab
gg_emean = gg_emean + theme_grey( base_size      = gg_ptsz
                                , base_family    = "Helvetica"
                                , base_line_size = 0.5
                                , base_rect_size = 0.5
                                )#end theme_grey

# Additional settings
gg_emean = gg_emean + theme( axis.text.x       = element_text( size   = gg_ptsz
                                                            , margin = unit(rep(0.35,times=4),"char")
                                                            )#end element_text
                           , axis.text.y       = element_text( size   = gg_ptsz
                                                            , margin = unit(rep(0.35,times=4),"char")
                                                            )#end element_text
                           , axis.ticks.length = unit(-0.2,"char")
                           , axis.title.y      = element_text( size = gg_ptsz * 0.9)
                           )#end theme

# Save plots.
for (d in sequence(ndevice)){
   e_output = paste0("emean_lai_mcd15a2h_v006.",gg_device[d])
   dummy    = ggsave( filename = e_output
                    , plot     = gg_emean
                    , device   = gg_device[d]
                    , path     = plot_path
                    , width    = gg_width
                    , height   = gg_height
                    , units    = gg_units
                    , dpi      = gg_depth
                    )#end ggsave

}#end for (d in sequence(ndevice))

# If sought, plot images on screen
if (gg_screen) plot(gg_emean)

```


```{r, label = 'plot-mmean'}
cat0(" + Plot the mean annual cycle of LAI for all sites.")

# Retrieve useful elements from the POI list
p_site   = poi_list$site
p_colour = poi_list$colour

# First, turn sites into factors
show_mmean = poi_mmean %>% mutate( site = factor(site,levels=poi_list$site))

# Plot time series with the range
gg_mmean = ggplot( data    = show_mmean
                 , mapping = aes_string( x      = "month"
                                       , y      = "lai_qmid"
                                       , ymin   = "lai_qlwr"
                                       , ymax   = "lai_qupr"
                                       , group  = "site"
                                       , colour = "site"
                                       , fill   = "site"
                                       )#end aes_string
                 )#end ggplot
gg_mmean = gg_mmean + scale_colour_manual(name="",aesthetics="colour",labels=p_site,values=p_colour)
gg_mmean = gg_mmean + scale_colour_manual(name="",aesthetics="fill"  ,labels=p_site,values=p_colour)

# Add ribbons with lower and upper range, and also the medians as lines
gg_mmean = gg_mmean + geom_ribbon( alpha = alpha_ribbon, colour = "transparent", show.legend=TRUE)
gg_mmean = gg_mmean + geom_line( lwd = 1.0, show.legend=TRUE)

# Add local annotation
gg_mmean = gg_mmean + labs(title=element_blank())
gg_mmean = gg_mmean + scale_x_continuous( breaks = sequence(12)
                                        , labels = substring(month.abb,1,1)
                                        )#end scale_x_continuous
gg_mmean = gg_mmean + xlab(element_blank())
gg_mmean = gg_mmean + ylab( desc.unit( desc = "MODIS Leaf Area Index (MCD15A2H)"
                                     , unit = untab$m2lom2
                                     , twolines = TRUE
                                     )#end desc.unit
                          )#end ylab
gg_mmean = gg_mmean + theme_grey( base_size      = gg_ptsz
                                , base_family    = "Helvetica"
                                , base_line_size = 0.5
                                , base_rect_size = 0.5
                                )#end theme_grey

# Additional settings
gg_mmean = gg_mmean + theme( axis.text.x       = element_text( size   = gg_ptsz
                                                            , margin = unit(rep(0.35,times=4),"char")
                                                            )#end element_text
                           , axis.text.y       = element_text( size   = gg_ptsz
                                                            , margin = unit(rep(0.35,times=4),"char")
                                                            )#end element_text
                           , axis.ticks.length = unit(-0.2,"char")
                           , axis.title.y      = element_text( size = gg_ptsz * 0.9)
                           )#end theme

# Save plots.
for (d in sequence(ndevice)){
   m_output = paste0("mmean_lai_mcd15a2h_v006.",gg_device[d])
   dummy    = ggsave( filename = m_output
                    , plot     = gg_mmean
                    , device   = gg_device[d]
                    , path     = plot_path
                    , width    = gg_width
                    , height   = gg_height
                    , units    = gg_units
                    , dpi      = gg_depth
                    )#end ggsave

}#end for (d in sequence(ndevice))

# If sought, plot images on screen
if (gg_screen) plot(gg_mmean)

```



Define the template for the global attributes.
```{r, label='global-attributes', message=FALSE, results='hide'}

# Define the code developer information (indirect way so the email is not visible).
developer_name  = c( 111L, 103L, 110L, 111L,  76L,  32L, 115L, 111L,  99L, 114L,  97L,  77L)
developer_email = c( 118L, 111L, 103L,  46L, 108L,  98L, 108L,  64L, 111L, 103L, 110L, 111L
                   , 108L, 109L)


# Define the template.  We will update the title in each time step.
att_template = list( title          = "To be replaced when looping through months"
                   , version        = dat_version
                   , date_created   = paste0(as.character(now(tzone="UTC")), "UTC")
                   , source_code    = "c6_modis-lai_poi.Rmd"
                   , code_notes     = "LAI (MCD15A2H) for benchmarking ELM-FATES and CLM-FATES"
                   , code_developer = paste0( intToUtf8(rev(developer_name))
                                            ," <"
                                            , intToUtf8(rev(developer_email))
                                            ,">"
                                            )#end paste0
                   , data_info      = paste0(" Data downloaded through AppEEARS"
                                            ," (https://lpdaac.usgs.gov/tools/appeears/)."
                                            ," MODIS (Collection 6) products used:"
                                            ," MCD15A2H (https://doi.org/10.5067/MODIS/MCD15A2H.006);"
                                            ," MCD12Q1 (https://doi.org/10.5067/MODIS/MCD12Q1.006)."
                                            )#end paste0
                   )#end list
```




Create a single NetCDF file with all the averages by month and year.  We do not save the monthly average across all years because simulations may or may not overlap with all the observation months and years.

```{r, label='write-output', message=FALSE, results='hide'}
for (p in sequence(n_poi_list)){
   # Useful short names
   p_site       = poi_list$site      [p]
   p_output     = poi_list$output    [p]
   p_lon        = poi_list$lon       [p]
   p_lat        = poi_list$lat       [p]
   p_dxy        = poi_list$dxy       [p]
   cat0(" + Generate output data for ",p_site,".")

   # Find precision for the coordinates (typically 1% of the grid size).
   p_outprec = -floor(log10(p_dxy*0.01))

   # Define output file
   mlai_base = paste0(p_output,"_c6modis-summ.nc")
   mlai_file = file.path(output_path,mlai_base)
  
   # Standardise coordinates (and make sure latitude cannot exceed the poles).
   outlon  = round(p_lon,p_outprec) %% 360.
   outlat  = max(-90,min(90-0.5*p_dxy,round(p_lat,p_outprec)))

   # Find edges.
   edge_w = (outlon - 0.5 * p_dxy) %% 360.
   edge_e = (outlon + 0.5 * p_dxy) %% 360.
   edge_s = outlat - 0.5 * p_dxy
   edge_n = outlat + 0.5 * p_dxy
   
   # Extract data for this POI.
   this_emean   = poi_emean                %>% 
                  filter(site %in% p_site) %>%
                  arrange(when)
   n_this_emean = nrow(this_emean)

   # Extract time, and turn it into a difference in months since the first time
   year_first  = year (this_emean$when[1])
   month_first = month(this_emean$when[1])
   time_first  = this_emean$when[1]

   # Extract time, and turn it into a difference in months
   tsince = as.numeric(difftime(this_emean$when,time_first,units="days"))
   
   # In case file exists, it will be re-created.
   cat0("   - Write averages by month and year to ",mlai_base,".")
   if (file.exists(mlai_file)) file.remove(mlai_file)
   
   # Add dimensions: longitude, latitude, and time. We do not automatically create the 
   # dimension variable for time because R would create it in double precision.  Instead,
   # we append variable time manually.
   xx  = ncdim_def( name="lon"   ,units="",vals=1L               ,create_dimvar=FALSE)
   yy  = ncdim_def( name="lat"   ,units="",vals=1L               ,create_dimvar=FALSE)
   tt  = ncdim_def( name="time"  ,units="",vals=seq_along(tsince),create_dimvar=FALSE)
   ss  = ncdim_def( name="scalar",units="",vals=1L               ,create_dimvar=FALSE)
   
   # List of dimensions, useful for setting variables.   
   nc_xy  = list   (xx,yy)
   nc_xyt = list(xx,yy,tt)
   nc_t   = list      (tt)
   nc_s   = list(ss)
   xy     = c(1,1)
   xyt    = c(1,1,n_this_emean)
   
   # Start list with variables. First we put the coordinates
   nc_vlist        = list()
   nc_vlist$LONGXY = ncvar_def( name     = "LONGXY"
                              , units    = "degrees_east"
                              , dim      = nc_xy
                              , missval  = undef_out
                              , longname = "longitude"
                              )#end ncvar_def
   nc_vlist$LATIXY = ncvar_def( name     = "LATIXY"
                              , units    = "degrees_north"
                              , dim      = nc_xy
                              , missval  = undef_out
                              , longname = "latitude"
                              )#end ncvar_def
   nc_vlist$EDGEW  = ncvar_def( name     = "EDGEW"
                              , units    = "degrees_east"
                              , dim      = nc_s
                              , missval  = undef_out
                              , longname = "western edge in atmospheric data"
                              )#end ncvar_def
   nc_vlist$EDGEE  = ncvar_def( name     = "EDGEE"
                              , units    = "degrees_east"
                              , dim      = nc_s
                              , missval  = undef_out
                              , longname = "eastern edge in atmospheric data"
                              )#end ncvar_def
   nc_vlist$EDGES  = ncvar_def( name     = "EDGES"
                              , units    = "degrees_north"
                              , dim      = nc_s
                              , missval  = undef_out
                              , longname = "southern edge in atmospheric data"
                              )#end ncvar_def
   nc_vlist$EDGEN  = ncvar_def( name     = "EDGEN"
                              , units    = "degrees_north"
                              , dim      = nc_s
                              , missval  = undef_out
                              , longname = "northern edge in atmospheric data"
                              )#end ncvar_def
   nc_vlist$time   = ncvar_def( name     = "time"
                              , units    = paste0( "days since ",as.character(time_first)
                                                 , " 00:00:00 UTC"
                                                 )#end paste0
                              , dim      = nc_t
                              , missval  = undef_out
                              , longname = "observation time"
                              )#end ncvar_def
   
   # Include LAI and fPAR.
   nc_vlist$LAI    = ncvar_def( name     = "LAI"
                              , units    = "m2/m2"
                              , dim      = nc_xyt
                              , missval  = undef_out
                              , longname = "Leaf area index"
                              )#end ncvar_def
   nc_vlist$FPAR   = ncvar_def( name     = "FPAR"
                              , units    = "1"
                              , dim      = nc_xyt
                              , missval  = undef_out
                              , longname = "Fraction of photosynthetically active radiation"
                              )#end ncvar_def

   # Create file
   nc_conn = nc_create(filename=mlai_file,vars=nc_vlist,verbose=FALSE)
   
   #---~---
   # Put coordinates, tower height and attributes to the netcdf
   #---~---
   # Longitude, append time-invariant tag
   dummy = ncvar_put(nc=nc_conn,varid="LONGXY",vals=array(data=outlon     ,dim=xy))
   dummy = ncatt_put(nc=nc_conn,varid="LONGXY",attname="mode"    ,attval="time-invariant")
   # Latitude, append time-invariant tag
   dummy = ncvar_put(nc=nc_conn,varid="LATIXY",vals=array(data=outlat     ,dim=xy))
   dummy = ncatt_put(nc=nc_conn,varid="LATIXY",attname="mode"    ,attval="time-invariant")
   # Western edge, append time-invariant tag
   dummy = ncvar_put(nc=nc_conn,varid="EDGEW" ,vals=edge_w)
   dummy = ncatt_put(nc=nc_conn,varid="EDGEW" ,attname="mode"    ,attval="time-invariant")
   # Eastern edge, append time-invariant tag
   dummy = ncvar_put(nc=nc_conn,varid="EDGEE" ,vals=edge_e)
   dummy = ncatt_put(nc=nc_conn,varid="EDGEE" ,attname="mode"    ,attval="time-invariant")
   # Southern edge, append time-invariant tag
   dummy = ncvar_put(nc=nc_conn,varid="EDGES" ,vals=edge_s)
   dummy = ncatt_put(nc=nc_conn,varid="EDGES" ,attname="mode"    ,attval="time-invariant")
   # Northern edge, append time-invariant tag
   dummy = ncvar_put(nc=nc_conn,varid="EDGEN" ,vals=edge_n)
   dummy = ncatt_put(nc=nc_conn,varid="EDGEN" ,attname="mode"    ,attval="time-invariant")
   # Time, append calendar type.
   dummy = ncvar_put(nc=nc_conn,varid="time"  ,vals=tsince)
   dummy = ncatt_put(nc=nc_conn,varid="time"  ,attname="calendar",attval="gregorian")
   # Put LAI to the netcdf
   dummy = ncvar_put(nc=nc_conn,varid="LAI"  ,vals=array(data=this_emean$lai_qmid,dim=xyt))
   dummy = ncatt_put(nc=nc_conn,varid="LAI"  ,attname="mode",attval="time-dependent")
   # Put FPAR to the netcdf
   dummy = ncvar_put(nc=nc_conn,varid="FPAR"  ,vals=array(data=this_emean$fpar_qmid,dim=xyt))
   dummy = ncatt_put(nc=nc_conn,varid="FPAR"  ,attname="mode",attval="time-dependent")

   # Add title specific for this month/year.
   nc_title   = paste0( "Averages by month and year for ",p_site)
   att_global = modifyList( x = att_template, val = list( title = nc_title ))
   
   
   # Loop through global attributes
   for (l in seq_along(att_global)){
      # Current attribute information
      att_name  = names(att_global)[l]
      att_value = att_global[[l]]
      
      # Add attribute 
      dummy = ncatt_put(nc=nc_conn,varid=0,attname=att_name,attval=att_value)
   }#end for (l in seq_along(att_global))
   
   
   # Close the file
   dummy = nc_close(nc_conn)
}#end for (p in sequence(n_poi_list))
```



