---
title: "FATES post-processing script"
author: "Marcos Longo"
date: "12-Jul-2021"
---

# Introduction

This R Markdown document guides how to build a set of output figures for a FATES simulation at a single 
site.  The script is (permanently?) under development, and new features may be incorporated over time.

To run this script, you will need to have FATES simulations, and a few additional scripts (typically located at the `<path to>/RUtils` directory.  Most scripts are generic utilities, but you may want to check and edit the following scripts:
* **fates_varlist.r**. This script contains a list of FATES-specific variables, which may be grouped by size class, age, PFT.
* **hlm_varlist.r**. This script contains a list of host land model variables, which may be useful for verifying the biogeophysical cycles.

Check the comments of these scripts for additional details.


# Reset session

Use this chunk to fully reset R.
```{r, label = 'reset-R',message=FALSE, results='hide'}
# Unload all packages except for the R default ones
plist = names(sessionInfo()$otherPkgs)
if (length(plist) > 0){
   dummy = sapply(X=paste0("package:",plist),FUN=detach,character.only=TRUE,unload=TRUE)
}#end if (length(plist) > 0)


# Remove all variables
rm(list=ls())

# Reset warnings
options(warn=0)

# Close all plots
invisible(graphics.off())

# Clean up
invisible(gc())
```





# Initial settings

First we set some useful paths and initial settings.

* `home.path`. The main path. Typically the main home/scratch directory where sub-directories are located.
* `hesm.path`. The main path to where the simulation outputs are written.
* `simul`. Label for the simulation to be plotted (typically also a sub-directory in `hesm.path`).
* `simul.desc`. Simulation description to be printed in plot titles. Not a path, but it is easier to forget setting this variable if it is not right next to `simul`.
* `simul.root`. List of all possible locations for output netCDF files. Normally both ELM and CLM paths should be included, so the script works with both.
* `srcdir`.  Directory with additional R scripts that have extra settings. **Important**. This must be downloaded from [GitHub](https://github.com/mpaiao/FATES_Utils).
* `plot.root`. Directory to where plots will be written.

```{r, label='path-settings',message=FALSE,results='hide'}
home.path  = path.expand("~")
hesm.root  = file.path(home.path,"Documents/LocalData/FATES/Simulations") # Current directory.
simul      = "D0005_ParacouTest_CLM_FATES"                     # File label for current simulation.
simul.desc = "D0005: Paracou test (CLM-FATES)"                 # Current simulation description.
simul.root = c( file.path(hesm.root,simul,"run")               # Main simulation path, list all
              , file.path(hesm.root,simul,"lnd","hist")        #   possibilities, the code will
              )#end c                                          #   use the first valid one.
srcdir     = file.path(home.path,"Util/RUtils")                # Path with additional scripts.
plot.root  = file.path(hesm.root,"figures",simul)              # Directory for figures.

# Output path for time series
tspoi.path    = file.path(plot.root,"tseries_poi"  )
tstheme.path  = file.path(plot.root,"tseries_theme")
tsage.path    = file.path(plot.root,"tseries_age"  )
tsdbh.path    = file.path(plot.root,"tseries_dbh"  )
tspft.path    = file.path(plot.root,"tseries_pft"  )

# Create output paths
dummy      = dir.create(tspoi.path  ,recursive = TRUE,showWarnings = FALSE)
dummy      = dir.create(tstheme.path,recursive = TRUE,showWarnings = FALSE)
dummy      = dir.create(tsage.path  ,recursive = TRUE,showWarnings = FALSE)
dummy      = dir.create(tsdbh.path  ,recursive = TRUE,showWarnings = FALSE)
dummy      = dir.create(tspft.path  ,recursive = TRUE,showWarnings = FALSE)


```


Set time range to be included in the plots.
```{r, label='set-times',message=FALSE,results='hide'}
whena      = "01/01/2004" # First time (MM/DD/YYYY)
whenz      = "12/01/2019" # Last time (MM/DD/YYYY)
```


_Optional_. Here we define the plant functional types to include in the plots. We define the basic PFT information 
in variable `pftinfo`, which is a list.  Each list element contains the following sub-elements:
* *id*. PFT index in FATES. 
* *key*. Short name to define PFT in data structures. Prefer strings that start with a letter, and have only letters, numbers, `.`, and `_`.
* **short**. Short name to define PFT in figure legends. Feel free to use spaces, `plotmath` symbols and other characters, but keep them short.
* **desc**. Long name to define PFTs, for general information.
* **colour**. Colour associated with PFT.  Prefer to use R's hexadecimal format (e.g. `"#1F78B4"`).

In some cases setting names and descriptions for PFTs will not be feasible (e.g., when many PFTs are included). In this case, set `user.pftinfo=FALSE`.

```{r, label='set-pfts',message=FALSE,results='hide'}
# Use the user-defined PFT settings?
user.pftinfo = TRUE

# List of PFTs, in case user.pftinfo is TRUE
n            = 0
pftinfo      = list()
n            = n + 1
pftinfo[[n]] = list( id     = 1
                   , key    = "pft01"
                   , short  = "GenBLTr"
                   , desc   = "Generic Broadleaf Tropical Tree"
                   , colour = "#66C2A4"
                   )#end list
```


General plot options for `ggplot`
```{r, label='set-ggplot',message=FALSE,results='hide'}
gg.device  = c("pdf") # Output devices to use (Check ggsave for acceptable formats)
gg.depth   = 300      # Plot resolution (dpi)
gg.ptsz    = 18       # Font size
gg.width   = 8.5      # Plot width (units below)
gg.height  = 8.5      # Plot height (units below)
gg.units   = "in"     # Units for plot size
gg.screen  = TRUE     # Show plots on screen as well?
gg.tfmt    = "%Y"     # Format for time 

ndevice = length(gg.device)
```

# Main script

__Note:__ Changes beyond this point are only needed if you are developing the notebook.

Load packages and tools
```{r, label='load-everything'}
source(file.path(srcdir,"load.everything.r"))
isok = require(ncdf4)
if (! isok) stop("This script requires package \"ncdf4\".")
```


Initial settings.
```{r, label='initial-settings',message=FALSE,results='hide'}

#--- Convert strings to dates.
whena  = chron(whena)
whenz  = chron(whenz)
montha = nummonths(whena)
yeara  = numyears(whena)
monthz = nummonths(whenz)
yearz  = numyears(whenz)
nyears = yearz-yeara
#---~---


#--- Build time series.
if (yeara == yearz){
  when = chron(paste(seq(from=montha,to=monthz,by=1)),1,yeara,sep="/")
}else{
  when = c( chron(paste(seq(from=montha,to=12,by=1),1,yeara,sep="/"))
          , chron(paste( rep(seq(from=1,to=12,by=1),times=nyears-1)
                       , 1
                       , rep(seq(from=yeara+1,to=yearz-1,by=1),each=12)
                       , sep="/"
                       )#end paste
                 )#end chron
          , chron(paste(seq(from=1,to=monthz,by=1),1,yearz,sep="/"))
          )#end c  
}#end if (yeara == yearz)
#---~---


#--- Number of times
nwhen = length(when)
#---~---

```


Use the first time to load a few settings (dimensions, soil information).
```{r, label='first-set',message=FALSE,results='hide'}
if ("nc.conn" %in% ls()){dummy = nc_close(nc.conn); rm(nc.conn)}

#--- Extract times and build all possible file names
w.month    = nummonths(when[1])
w.year     = numyears (when[1])
w.ymlab    = sprintf("%4.4i-%2.2i",w.year,w.month)
hlm.midfix = c("elm.h0","clm2.h0")
nc.base    = paste0(simul,".",hlm.midfix,"." ,w.ymlab,".nc")
nc.file    = file.path( rep(x = simul.root, times = length(nc.base   ))
                      , rep(x = nc.base   , each  = length(simul.root))
                      )#end file.path
#---~---
  

  
#--- Identify which model and path to use.
idx = which(file.exists(nc.file))
if (length(idx) > 0){
   #--- Load the correct file.
   nc.file=(nc.file[idx])[1]
   nc.base=basename(nc.file)
   cat0(" + Read data from file ",nc.base,".")
   #---~---

   #--- Remember which path and model to use for other times
   simul.root = dirname(nc.file)
   sel.midfix = mapply(FUN=grepl,pattern=as.list(hlm.midfix),MoreArgs=list(x=nc.base))
   hlm.midfix = hlm.midfix[sel.midfix]
   #---~---
}else{
   #--- No file was found. Stop the run
   cat0(" + Neither CLM nor ELM expected files found.")
   cat0("   - Expected ELM file:    ",nc.file[1],".")
   cat0("   - Expected CLM file:    ",nc.file[4],".")
   cat0("   - Alternative ELM file: ",nc.file[2],".")
   cat0("   - Alternative CLM file: ",nc.file[3],".")
   stop(" Invalid files, perhaps incorrect simulation?")
   #---~---
}#end if (length(idx) > 0)
#---~---
  
  
#--- Open NetCDF connection and retrieve variable names
nc.conn  = nc_open(filename=nc.file)
nc.nvars = nc.conn$nvars
nc.ndims = nc.conn$ndims
nc.dlist = rep(NA_character_,times=nc.ndims)
nc.vlist = rep(NA_character_,times=nc.nvars)
for (d in sequence(nc.ndims)) nc.dlist[d] = nc.conn$dim[[d]]$name
for (v in sequence(nc.nvars)) nc.vlist[v] = nc.conn$var[[v]]$name
#---~---

#---~---
# Gather dimension information, then initialise matrices
#---~---

#--- List of age classes
idxage   = match("fates_levage",nc.dlist)
if (is.finite(idxage)){
   ages     = nc.conn$dim[[idxage]]$vals
   nages    = nc.conn$dim[[idxage]]$len
   agekeys  = sprintf("age.%3.3i",ages)
   agelabs  = c( paste0("paste(",ages[-nages],"-",ages[-1],")")
               , paste0("paste(",ages[nages],"-infinity)")
               )#end c
   agelabs  = parse(text=agelabs)
}else{
   ages     = numeric(0L)
   nages    = 0L
   agekeys  = character(0L)
   agelabs  = character(0L)
}#end if (is.na(idxage))
#---~---

#--- List of size classes
idxdbh   = match("fates_levscls",nc.dlist)
if (is.finite(idxdbh)){
   dbhs     = nc.conn$dim[[idxdbh]]$vals
   ndbhs    = nc.conn$dim[[idxdbh]]$len
   dbhkeys  = sprintf("dbh.%3.3i",dbhs)
   dbhlabs  = c( paste0("paste(",dbhs[-ndbhs],"-",dbhs[-1],")")
               , paste0("paste(",dbhs[ndbhs],"-infinity)")
               )#end dbhlabs
   dbhlabs  = parse(text=dbhlabs)
}else{
   dbhs    = numeric(0L)
   ndbhs   = 0L
   dbhkeys = character(0L)
   dbhlabs = character(0L)
}#end if (is.finite(idxdbh))
#---~---
 
#--- List of PFT classes (only if not using user-defined classes).
idxpft   = match("fates_levpft",nc.dlist)
if (! is.finite(idxpft)){
   #--- PFT index not found. Skip PFTs altogether.
   pftinfo = data.table( id               = numeric(0L)
                       , key              = character(0L)
                       , short            = character(0L)
                       , desc             = character(0L)
                       , colour           = character(0L)
                       , stringsAsFactors = FALSE
                       )#end data.table
   #---~---
}else if (! user.pftinfo){
   #--- Select all PFTs available
   pftids  = nc.conn$dim[[idxpft]]$vals
   npftids = nc.conn$dim[[idxpft]]$len
   #---~---

   #--- Build data table with all the PFTs.
   pftinfo = data.table( id               = pftids
                       , key              = sprintf("pft%2.2i" ,pftids)
                       , short            = sprintf("PFT%2.2i" ,pftids)
                       , desc             = sprintf("PFT %2.2i",pftids)
                       , colour           = brewer.pal(n=npftids,name="PuBuGn")
                       , stringsAsFactors = FALSE
                       )#end data.table
   #---~---
}else{
   #--- Convert user-defined pftinfo to a "data table" object
   pftinfo  = do.call(what=rbind,args=lapply(X=pftinfo,FUN=as.data.table,stringsAsFactors=FALSE))
}#end if (! is.finite(idxpft))
#---~---

#--- Set number of PFTs (active PFTs only) 
npfts = nrow(pftinfo)
#---~---
   
#--- Retrieve all variables by age class. We also test whether
nc.byage = nc.vlist[grepl(pattern="_BY_AGE$",x=nc.vlist)]
nc.pref  = tolower(gsub(pattern="_BY_AGE$",replacement="",x=nc.byage))
nc.keep  = nc.pref %in% fatesvar$vnam
no.byage = nc.byage[! nc.keep]
nc.byage = nc.byage[  nc.keep]
nbyage   = length(nc.byage)
#---~---

#---~---
#    Retrieve all variables by size class. We also test whether total LAI can be
# obtained from under storey and canopy.
#---~---
is.size   = grepl(pattern="_SCLS$",x=nc.vlist) | grepl(pattern="_SCPF$",x=nc.vlist)
nc.bydbh  = nc.vlist[is.size]
nc.pref   = gsub(pattern="_SCLS$",replacement="",x=nc.bydbh)
nc.pref   = gsub(pattern="_SCPF$",replacement="",x=nc.pref )
nc.pref   = tolower(nc.pref)
nc.keep   = nc.pref %in% fatesvar$vnam
no.bydbh  = nc.bydbh[! nc.keep]
nc.bydbh  = unique(nc.bydbh[  nc.keep])
if (  all(c("LAI_UNDERSTORY_SCLS","LAI_CANOPY_SCLS") %in% nc.bydbh)
   && (! "LAI_SCLS" %in% nc.bydbh) ){
   nc.bydbh    = unique(c(nc.bydbh,"LAI_SCLS"))
   laidbh.last = TRUE
}else if (  all(c("LAI_UNDERSTORY_SCPF","LAI_CANOPY_SCPF") %in% nc.bydbh)
   && (! "LAI_SCPF" %in% nc.bydbh) ){
   nc.bydbh    = unique(c(nc.bydbh,"LAI_SCPF"))
   laidbh.last = TRUE
}else{
   laidbh.last = FALSE
}#end if (  all(c("LAI_UNDERSTORY_SCLS","LAI_CANOPY_SCLS") %in% nc.bydbh)
nbydbh    = length(nc.bydbh)
#---~---

    
#---~---
#    Retrieve all variables by PFT. We also test whether total LAI can be
# obtained from under storey and canopy.
#---~---
is.pft    = grepl(pattern="_SCPF$",x=nc.vlist)
nc.bypft  = nc.vlist[is.pft]
nc.pref   = tolower(gsub(pattern="_SCPF$",replacement="",x=nc.bypft ))
nc.keep   = nc.pref %in% fatesvar$vnam
no.bypft  = nc.bypft[! nc.keep]
nc.bypft  = unique(nc.bypft[  nc.keep])
if (  all(c("LAI_UNDERSTORY_SCPF","LAI_CANOPY_SCPF") %in% nc.bypft)
   && (! "LAI_SCPF" %in% nc.bypft) ){
   nc.bypft    = unique(c(nc.bypft,"LAI_SCPF"))
   laipft.last = TRUE
}else{
   laipft.last = FALSE
}#end if (  all(c("LAI_UNDERSTORY_SCPF","LAI_CANOPY_SCPF") %in% nc.bypft)
nbypft    = length(nc.bypft)
#---~---


#--- Retrieve all "1D" variables that are available at the host model.
nc.pref   = tolower(x=nc.vlist)
nc.keep   = nc.pref %in% hlm1dvar$vnam
no.hlm1d  = nc.vlist[! nc.keep]
nc.hlm1d  = nc.vlist[  nc.keep]
if (  ( all(c("QSOIL","QVEGT","QVEGE") %in% nc.bydbh) )
   && (! "QEVTR" %in% nc.bydbh) ){
   nc.hlm1d = unique(c(nc.bydbh,"QEVTR"))
   etr.last = TRUE
}else{
   etr.last = FALSE
}#end if (all(c("QSOIL","QVEGT","QVEGE") %in% nc.bydbh))
nhlm1d    = length(nc.hlm1d)
#---~---


#--- Initialise list of variables by age class.
byage  = list()
for (a in sequence(nbyage)){
   nc.nvnow        = nc.byage[a]
   nc.pref         = tolower(gsub(pattern="_BY_AGE$",replacement="",x=nc.nvnow))
   f               = match(nc.pref,fatesvar$vnam)
   f.vnam          = fatesvar$vnam[f]
   byage[[f.vnam]] = matrix(data=NA_real_,nrow=nwhen,ncol=nages,dimnames=list(NULL,agekeys))
}#end for (a in sequence(nbyage))
#---~---

#--- Initialise list of variables by size class
bydbh = list()
for (d in sequence(nbydbh)){
   nc.nvnow        = nc.bydbh[d]
   nc.pref         = gsub(pattern="_SCLS$",replacement="",x=nc.nvnow)
   nc.pref         = gsub(pattern="_SCPF$",replacement="",x=nc.pref )
   nc.pref         = tolower(nc.pref)
   f               = match(nc.pref,fatesvar$vnam)
   f.vnam          = fatesvar$vnam[f]
   bydbh[[f.vnam]] = matrix(data=NA_real_,nrow=nwhen,ncol=ndbhs,dimnames=list(NULL,dbhkeys))
}#end for (d in sequence(nbydbh))
#---~---

#--- Initialise list of variables by PFT class
bypft = list()
for (p in sequence(nbypft)){
   nc.nvnow        = nc.bypft[p]
   nc.pref         = tolower(gsub(pattern="_SCPF$",replacement="",x=nc.nvnow))
   nc.pref         = tolower(nc.pref)
   f               = match(nc.pref,fatesvar$vnam)
   f.vnam          = fatesvar$vnam[f]
   bypft[[f.vnam]] = matrix(data=NA_real_,nrow=nwhen,ncol=npfts,dimnames=list(NULL,pftinfo$key))
}#end for (d in sequence(nbypft))
#---~---

#--- Initialise 1D variables available at the HLM
hlm1d = as.data.table( matrix( data     = NA_real_
                             , nrow     = nwhen
                             , ncol     = nhlm1d
                             , dimnames = list(NULL,tolower(nc.hlm1d))
                             )#end matrix
                     )#end as.data.table
#---~---


#--- Load soil layers
slayer = data.table( zsoi   = ncvar_get(nc=nc.conn,varid='ZSOI'  )
                   , dzsoi  = ncvar_get(nc=nc.conn,varid='DZSOI' )
                   , bsw    = ncvar_get(nc=nc.conn,varid='BSW'   )
                   , hksat  = ncvar_get(nc=nc.conn,varid='HKSAT' )
                   , sucsat = ncvar_get(nc=nc.conn,varid='SUCSAT')
                   , watsat = ncvar_get(nc=nc.conn,varid='WATSAT')
                   )#end data.table
#---~---


#--- Load indices
index_scpf = data.table( scls   = ncvar_get(nc=nc.conn,varid='fates_scmap_levscpf'  )
                       , pft    = ncvar_get(nc=nc.conn,varid='fates_pftmap_levscpf' )
                       )#end data.table
#---~---

# Close connection
dummy   = nc_close(nc.conn)

```


Loop through the times, and load the data sets.
```{r, label='load-times', message=FALSE}
if ("nc.conn" %in% ls()){dummy = nc_close(nc.conn); rm(nc.conn)}
for (w in sequence(nwhen)){
   #--- Extract times and build file name
   w.month  = nummonths(when[w])
   w.year   = numyears (when[w])
   w.ymlab  = sprintf("%4.4i-%2.2i",w.year,w.month)
   nc.base  = paste0(simul,".",hlm.midfix,"." ,w.ymlab,".nc")
   nc.file  = file.path(simul.root,nc.base)
   #---~---

   #-- Find conversion factors for monthly variables.
   cmon.day = daymax(w.month,w.year)
   cmon.hr  = day.hr  * cmon.day
   cmon.min = day.min * cmon.day
   cmon.sec = day.sec * cmon.day
   #---~---

  
   #--- Open NetCDF connection and retrieve variable names
   nc.conn  = nc_open(filename=nc.file)
   nc.nvars = nc.conn$nvars
   nc.vlist = rep(NA_character_,times=nc.nvars)
   for (v in sequence(nc.nvars)) nc.vlist[v] = nc.conn$var[[v]]$name
   #---~---


    

   #--- Read variables by age, and assign current values to the matrix.  
   for (a in sequence(nbyage)){
      nc.nvnow            = nc.byage[a]
      nc.pref             = tolower(gsub(pattern="_BY_AGE$",replacement="",x=nc.nvnow))
      f                   = match(nc.pref,fatesvar$vnam)
      f.vnam              = fatesvar$vnam[f]
      f.add0              = eval(parse(text=fatesvar$add0[f]))
      f.mult              = eval(parse(text=fatesvar$mult[f]))
      nc.dat              = ncvar_get(nc=nc.conn,varid=nc.nvnow)
      byage[[f.vnam]][w,] = f.add0 + f.mult * nc.dat
   }#end for (a in sequence(nbyage))
   #---~---

   #---~--- 
   #   Read variables by size, and assign current values to the matrix.  In case LAI
   # is in the list and it is the last variable, we calculate it from canopy and under
   # storey, after loading all variables
   #---~---
   for (d in sequence(nbydbh-laidbh.last)){
      nc.nvnow            = nc.bydbh[d]
      is.scpf             = grepl(pattern="_SCPF$",x=nc.nvnow)
      nc.pref             = gsub(pattern="_SCLS$",replacement="",x=nc.nvnow)
      nc.pref             = gsub(pattern="_SCPF$",replacement="",x=nc.pref )
      nc.pref             = tolower(nc.pref)
      f                   = match(nc.pref,fatesvar$vnam)
      f.vnam              = fatesvar$vnam[f]
      f.add0              = eval(parse(text=fatesvar$add0[f]))
      f.mult              = eval(parse(text=fatesvar$mult[f]))
      f.aggr              = match.fun(fatesvar$aggr[f])
      nc.dat              = ncvar_get(nc=nc.conn,varid=nc.nvnow)
      nc.dat              = f.add0 + f.mult * nc.dat
      if (is.scpf){
         #--- Aggregate data by size class
         nc.aggr        = tapply( X     = nc.dat
                                , INDEX = index_scpf$scls
                                , FUN   = f.aggr
                                , na.rm = TRUE
                                )#end tapply
         names(nc.aggr) = NULL
         bydbh[[f.vnam]][w,] = nc.aggr
         #---~---
      }else{
         #--- Variable is truly a size class.
         bydbh[[f.vnam]][w,] = nc.dat
         #---~---
      }#end if (is.scpf)
      #---~---
   }#end for (d in sequence(nbydbh-laidbh.last))
   #---~---

   #--- Find total LAI if sought.
   if (laidbh.last){
      bydbh$lai[w,] = bydbh$lai_canopy[w,] + bydbh$lai_understory[w,]
   }#end if (lai.last)
   #---~---



   #---~--- 
   #   Read variables by PFT, and assign current values to the matrix.  In case LAI
   # is in the list and it is the last variable, we calculate it from canopy and under
   # storey, after loading all variables
   #---~---
   for (p in sequence(nbypft-laipft.last)){
      # Load variable information
      nc.nvnow = nc.bypft[p]
      nc.pref  = tolower(gsub(pattern="_SCPF$",replacement="",x=nc.nvnow))
      f        = match(nc.pref,fatesvar$vnam)
      f.vnam   = fatesvar$vnam[f]
      f.add0   = eval(parse(text=fatesvar$add0[f]))
      f.mult   = eval(parse(text=fatesvar$mult[f]))
      f.aggr   = match.fun(fatesvar$aggr[f])
      f.dbh01  = fatesvar$dbh01[f]
      #---~---

      #--- Retrieve data.
      nc.dat   = ncvar_get(nc=nc.conn,varid=nc.nvnow)
      #---~---

      #--- Decide whether or not to exclude the first DBH class.
      if (f.dbh01){
         seldbh = rep(TRUE,times=length(nc.dat))
      }else{
         seldbh = ! (index_scpf$scls %in% c(1))
      }#end if (f.dbh01)
      #---~---

      #--- Aggregate data by size class
      nc.aggr        = tapply( X     = f.add0 + f.mult * nc.dat[seldbh]
                             , INDEX = index_scpf$pft[seldbh]
                             , FUN   = f.aggr
                             , na.rm = TRUE
                             )#end tapply
      names(nc.aggr) = NULL
      #---~---

      #--- Brind only the PFTs we are interested in.
      bypft[[f.vnam]][w,] = nc.aggr[pftinfo$id]
      #---~---
   }#end for (d in sequence(nbydbh-laidbh.last))
   #---~---

   #--- Find total LAI if sought.
   if (laipft.last){
      bypft$lai[w,] = bypft$lai_canopy[w,] + bypft$lai_understory[w,]
   }#end if (laipft.last)
   #---~---


    
   #--- Read 1D variables  
   for (v in sequence(nhlm1d-etr.last)){
      nc.nvnow            = nc.hlm1d[v]
      nc.pref             = tolower(x=nc.nvnow)
      h                   = match(nc.pref,hlm1dvar$vnam)
      h.vnam              = hlm1dvar$vnam[h]
      h.add0              = eval(parse(text=hlm1dvar$add0[h]))
      h.mult              = eval(parse(text=hlm1dvar$mult[h]))
      nc.dat              = ncvar_get(nc=nc.conn,varid=nc.nvnow)
      hlm1d[[h.vnam]][w]  = h.add0 + h.mult * nc.dat
   }#for (h in sequence(nhlm1d))
   #---~---

   #--- Find total ET.
   if (etr.last){
      bydbh$qevtr[w,] = bydbh$qeve[w,] + bydbh$qevtr[w,] + bydbh$qsoil[w,]
   }#end if (lai.last)
   #---~---

   # Close connection
   dummy   = nc_close(nc.conn)
  
}#end for (w in sequence(nwhen))




```

Scale variables by age that should be reported as stacks with the patch area by age. 
Note that patch area by age must be always included in FATES output. Patch area by age itself should not be scaled.
```{r, label='scale-age'}
#--- Scale patch variables by age that should be stacked. 
cat0(" + Rescale variables by age class.")
for (a in sequence(nbyage)){
   #--- Match variables
   f        = match(names(byage)[a],fatesvar$vnam)
   f.vnam   = fatesvar$vnam [f]
   f.desc   = fatesvar$desc [f]
   f.stack  = fatesvar$stack[f]
   #---~---

   #--- Proceed only if the variable is stacked and not the patch area.
   if (f.stack && (! (f.vnam %in% "patch_area"))){
      cat0("   - ",f.desc,".")
      byage[[f.vnam]] = byage[[f.vnam]] * byage$patch_area
   }#end if (f.stack && (! (f.vnam %in% "patch_area")))
      
}#end for (a in sequence(nbyage))

```


Turn data matrices into molten data tables (preferred data structure for plotting and analysing).
```{r, label='melt-data'}

#--- Turn age-dependent matrices into data tables
cat0(" + Turn age-dependent matrices into data tables.")
age.melt = NULL
for (a in sequence(nbyage)){
   #--- Match variables.
   f        = match(names(byage)[a],fatesvar$vnam)
   f.vnam   = fatesvar$vnam[f]
   f.desc   = fatesvar$desc[f]
   cat0("   - ",f.desc,".")
   #---~---

   #--- Create molten data table for this variable.  
   now.age      = as.data.table(byage[[f.vnam]])
   now.age$when = chron(when)
   now.melt     = data.table(melt(data=now.age,id.vars="when",variable.name="age",value.name=f.vnam))
   #---~---

   #--- Merge data table
   if (is.null(age.melt)){
      age.melt = now.melt
   }else{
      age.melt = merge(x=age.melt,y=now.melt,by=c("when","age"),all=TRUE)
   }#end if (is.null(age.melt))
   #---~---
}#end for (a in sequence(nbyage))
#---~---

#--- Turn size-dependent matrices into data tables
cat0(" + Turn size-dependent matrices into data tables.")
dbh.melt = NULL
for (d in sequence(nbydbh)){
   #--- Match variables.
   f        = match(names(bydbh)[d],fatesvar$vnam)
   f.vnam   = fatesvar$vnam[f]
   f.desc   = fatesvar$desc[f]
   cat0("   - ",f.desc,".")
   #---~---

   #--- Create molten data table for this variable.  
   now.dbh      = as.data.table(bydbh[[d]])
   now.dbh$when = chron(when)
   now.melt     = data.table(melt(data=now.dbh,id.vars="when",variable.name="dbh",value.name=f.vnam))
   #---~---

   #--- Merge data table
   if (is.null(dbh.melt)){
      dbh.melt = now.melt
   }else{
      dbh.melt = merge(x=dbh.melt,y=now.melt,by=c("when","dbh"),all=TRUE)
   }#end if (is.null(scls.melt))
   #---~---
}#end for (d in sequence(nbydbh))
#---~---

#--- Turn PFT-dependent matrices into data tables
cat0(" + Turn PFT-dependent matrices into data tables.")
pft.melt = NULL
for (p in sequence(nbypft)){
   #--- Match variables.
   f        = match(names(bypft)[p],fatesvar$vnam)
   f.vnam   = fatesvar$vnam[f]
   f.desc   = fatesvar$desc[f]
   cat0("   - ",f.desc,".")
   #---~---

   #--- Create molten data table for this variable.  
   now.pft      = as.data.table(bypft[[p]])
   now.pft$when = chron(when)
   now.melt     = data.table(melt(data=now.pft,id.vars="when",variable.name="pft",value.name=f.vnam))
   #---~---

   #--- Merge data table
   if (is.null(pft.melt)){
      pft.melt = now.melt
   }else{
      pft.melt = merge(x=pft.melt,y=now.melt,by=c("when","pft"),all=TRUE)
   }#end if (is.null(pft.melt))
   #---~---
}#end for (p in sequence(nbypft))
#---~---

#--- Remove data that are not molten.
byage = age.melt
bydbh = dbh.melt
bypft = pft.melt
rm(age.melt,dbh.melt,pft.melt)
#---~---

#--- Convert dbh and age classes into integers
cat0(" + Convert classes to integers.")
byage$age = as.numeric(byage$age)
bydbh$dbh = as.numeric(bydbh$dbh)
bypft$pft = as.numeric(bypft$pft)
#---~---
```



Plot time series by age:
```{r, label='plot-tsage',message=FALSE,results='hide'}
cat0(" + Plot time series of age-dependent variables.")

#--- Title for legend
age.legend = desc.unit(desc="Age",unit=untab$yr)
#---~---

age.loop   = which(fatesvar$vnam %in% names(byage))
gg.age   = list()
for (f in age.loop){
  #--- Match variables.
  f.vnam   = fatesvar$vnam [f]
  f.desc   = fatesvar$desc [f]
  f.unit   = fatesvar$unit [f]
  f.stack  = fatesvar$stack[f]
  cat0("   - ",f.desc,".")
  #---~---
  
  #--- Temporary data table. We convert the classes back to factor.
  f.byage     = byage
  f.byage$age = factor(f.byage$age,levels=sequence(nages))
  f.colages   = viridis(nages,option="D",direction=-1)
  f.agelabs   = agelabs
  #---~---
  
  #--- Initialise plot (decide whether to plot lines or stacks).
  if (f.stack){
     gg.now = ggplot(data=f.byage,aes_string(x="when",y=f.vnam,group="age",fill="age"))
     gg.now = gg.now + scale_fill_manual(name=age.legend,labels=f.agelabs,values=f.colages)
     gg.now = gg.now + geom_area(position="stack",show.legend = TRUE)
  }else{
     gg.now = ggplot(data=f.byage,aes_string(x="when",y=f.vnam,group="age",colour="age"))
     gg.now = gg.now + scale_colour_manual(name=age.legend,labels=f.agelabs,values=f.colages)
     gg.now = gg.now + geom_line(lwd=1.0,show.legend = TRUE)
  }#end if (f.stack)
  gg.now = gg.now + labs(title=simul.desc)
  gg.now = gg.now + scale_x_chron(format=gg.tfmt)
  gg.now = gg.now + xlab("Simulation time")
  gg.now = gg.now + ylab(desc.unit(desc=f.desc,unit=untab[[f.unit]],twolines=TRUE))
  gg.now = gg.now + theme_grey( base_size = gg.ptsz, base_family = "Helvetica",base_line_size = 0.5,base_rect_size =0.5)
  gg.now = gg.now + theme( legend.position   = "bottom"
                         , axis.text.x       = element_text( size   = gg.ptsz
                                                           , margin = unit(rep(0.35,times=4),"cm")
                                                           )#end element_text
                         , axis.text.y       = element_text( size   = gg.ptsz
                                                           , margin = unit(rep(0.35,times=4),"cm")
                                                           )#end element_text
                         , axis.ticks.length = unit(-0.25,"cm")
                         )#end theme
  #---~---
  
  #--- Save plot.
  for (d in sequence(ndevice)){
    f.output = paste0(f.vnam,"-tsage-",simul,".",gg.device[d])
    dummy = ggsave( filename = f.output
                  , plot     = gg.now
                  , device   = gg.device[d]
                  , path     = tsage.path
                  , width    = gg.width
                  , height   = gg.height
                  , units    = gg.units
                  , dpi      = gg.depth
                  )#end ggsave
  }#end for (o in sequence(nout))
  #---~---

  #--- Write plot settings to the list.
  gg.age[[f.vnam]] = gg.now
  #---~---
}#end for (a in age.loop)


#--- If sought, plot images on screen
if (gg.screen) gg.age
#---~---

```



Plot time series by size class:
```{r, label='plot-tsdbh',message=FALSE,results='hide'}
cat0(" + Plot time series of size-dependent variables.")

#--- Title for legend
dbh.legend = desc.unit(desc="DBH",unit=untab$cm)
#---~---

dbh.loop = which(fatesvar$vnam %in% names(bydbh))
gg.dbh   = list()
for (f in dbh.loop){
  #--- Match variables.
  f.vnam   = fatesvar$vnam [f]
  f.desc   = fatesvar$desc [f]
  f.unit   = fatesvar$unit [f]
  f.stack  = fatesvar$stack[f]
  f.dbh01  = fatesvar$dbh01[f]
  cat0("   - ",f.desc,".")
  #---~---
  
  #--- Decide whether to plot the first class.
  if (f.dbh01){
    #--- Keep all classes.
    f.bydbh     = bydbh
    f.bydbh$dbh = factor(f.bydbh$dbh,levels=sequence(ndbhs))
    f.coldbhs   = magma(ndbhs,direction=1)
    f.dbhlabs   = dbhlabs
    #---~---
  }else{
    #--- Exclude first class.
    bye         = as.numeric(bydbh$dbh) %in% 1
    f.bydbh     = bydbh[! bye,]
    f.bydbh$dbh = factor(f.bydbh$dbh,levels=sequence(ndbhs)[-1])
    f.coldbhs   = magma(ndbhs,direction=1)[-1]
    f.dbhlabs   = dbhlabs[-1]
    #---~---
  }#end if (f.dbh01)
  #---~---
   
  #--- Initialise plot (decide whether to plot lines or stacks).
  if (f.stack){
     gg.now = ggplot(data=f.bydbh,aes_string(x="when",y=f.vnam,group="dbh",fill="dbh"))
     gg.now = gg.now + scale_fill_manual(name=dbh.legend,labels=f.dbhlabs,values=f.coldbhs)
     gg.now = gg.now + geom_area(position=position_stack(reverse = TRUE),show.legend = TRUE)
  }else{
     gg.now = ggplot(data=f.bydbh,aes_string(x="when",y=f.vnam,group="dbh",colour="dbh"))
     gg.now = gg.now + scale_colour_manual(name=dbh.legend,labels=f.dbhlabs,values=f.coldbhs)
     gg.now = gg.now + geom_line(lwd=1.0,show.legend = TRUE)
  }#end if (f.stack)
  gg.now = gg.now + labs(title=simul.desc)
  gg.now = gg.now + scale_x_chron(format=gg.tfmt)
  gg.now = gg.now + xlab("Simulation time")
  gg.now = gg.now + ylab(desc.unit(desc=f.desc,unit=untab[[f.unit]],twolines=TRUE))
  gg.now = gg.now + theme_grey( base_size = gg.ptsz, base_family = "Helvetica",base_line_size = 0.5,base_rect_size =0.5)
  gg.now = gg.now + theme( legend.position   = "bottom"
                         , axis.text.x       = element_text( size   = gg.ptsz
                                                           , margin = unit(rep(0.35,times=4),"cm")
                                                           )#end element_text
                         , axis.text.y       = element_text( size   = gg.ptsz
                                                           , margin = unit(rep(0.35,times=4),"cm")
                                                           )#end element_text
                         , axis.ticks.length = unit(-0.25,"cm")
                         )#end theme
  #---~---

  #--- Save plot.
  for (d in sequence(ndevice)){
    f.output = paste0(f.vnam,"-tsdbh-",simul,".",gg.device[d])
    dummy = ggsave( filename = f.output
                  , plot     = gg.now
                  , device   = gg.device[d]
                  , path     = tsdbh.path
                  , width    = gg.width
                  , height   = gg.height
                  , units    = gg.units
                  , dpi      = gg.depth
                  )#end ggsave
  }#end for (o in sequence(nout))
  #---~---

  #--- Write plot settings to the list.
  gg.dbh[[f.vnam]] = gg.now
  #---~---
}#end for (a in age.loop)


#--- If sought, plot images on screen
if (gg.screen) gg.dbh
#---~---

```



Plot time series by plant functional type:
```{r, label='plot-tspft',message=FALSE,results='hide'}
cat0(" + Plot time series of size-dependent variables.")

#--- Title for legend
pft.legend = "Plant functional types"
#---~---

pft.loop = which(fatesvar$vnam %in% names(bypft))
gg.pft   = list()
for (f in pft.loop){
  #--- Match variables.
  f.vnam   = fatesvar$vnam [f]
  f.desc   = fatesvar$desc [f]
  f.unit   = fatesvar$unit [f]
  f.stack  = fatesvar$stack[f]
  cat0("   - ",f.desc,".")
  #---~---
  
  #--- Set plotting characteristics.
  f.bypft     = bypft
  f.bypft$pft = factor(f.bypft$pft,levels=sequence(npfts))
  f.colpfts   = pftinfo$colour
  f.pftlabs   = pftinfo$short
  #---~---

  #--- Initialise plot (decide whether to plot lines or stacks).
  if (f.stack){
     gg.now = ggplot(data=f.bypft,aes_string(x="when",y=f.vnam,group="pft",fill="pft"))
     gg.now = gg.now + scale_fill_manual(name=pft.legend,labels=f.pftlabs,values=f.colpfts)
     gg.now = gg.now + geom_area(position=position_stack(reverse = FALSE),show.legend = TRUE)
  }else{
     gg.now = ggplot(data=f.bypft,aes_string(x="when",y=f.vnam,group="pft",colour="pft"))
     gg.now = gg.now + scale_colour_manual(name=pft.legend,labels=f.pftlabs,values=f.colpfts)
     gg.now = gg.now + geom_line(lwd=1.0,show.legend = TRUE)
  }#end if (f.stack)
  gg.now = gg.now + labs(title=simul.desc)
  gg.now = gg.now + scale_x_chron(format=gg.tfmt)
  gg.now = gg.now + xlab("Simulation time")
  gg.now = gg.now + ylab(desc.unit(desc=f.desc,unit=untab[[f.unit]],twolines=TRUE))
  gg.now = gg.now + theme_grey( base_size = gg.ptsz, base_family = "Helvetica",base_line_size = 0.5,base_rect_size =0.5)
  gg.now = gg.now + theme( legend.position   = "bottom"
                         , axis.text.x       = element_text( size   = gg.ptsz
                                                           , margin = unit(rep(0.35,times=4),"cm")
                                                           )#end element_text
                         , axis.text.y       = element_text( size   = gg.ptsz
                                                           , margin = unit(rep(0.35,times=4),"cm")
                                                           )#end element_text
                         , axis.ticks.length = unit(-0.25,"cm")
                         )#end theme
  #---~---

  #--- Save plot.
  for (d in sequence(ndevice)){
    f.output = paste0(f.vnam,"-tspft-",simul,".",gg.device[d])
    dummy = ggsave( filename = f.output
                  , plot     = gg.now
                  , device   = gg.device[d]
                  , path     = tspft.path
                  , width    = gg.width
                  , height   = gg.height
                  , units    = gg.units
                  , dpi      = gg.depth
                  )#end ggsave
  }#end for (o in sequence(nout))
  #---~---

  #--- Write plot settings to the list.
  gg.pft[[f.vnam]] = gg.now
  #---~---
}#end for (a in age.loop)


#--- If sought, plot images on screen
if (gg.screen) gg.pft
#---~---

```



Plot theme time series:
```{r, label='plot-tstheme',message=FALSE,results='hide'}
cat0(" + Plot time series of thematically linked variables.")

gg.theme   = list()
for (th in sequence(ntstheme)){
  #--- Match variables.
  th.thnam  = tstheme$thnam  [th]
  th.thdesc = tstheme$thdesc [th]
  th.thunit = tstheme$thunit [th]
  th.vnames = tstheme$vnames [th]
  th.vcols  = tstheme$vcols  [th]
  th.stack  = tstheme$tsstack[th]

  #--- Split variables to include in this plot (plot only when all data are available).
  th.vlist   = c(unlist(strsplit(x=th.vnames,split="\\+")))
  th.match   = match(th.vlist,hlm1dvar$vnam)
  th.vdesc   = hlm1dvar$desc[th.match]
  th.labels  = parse(text=paste0("paste(",hlm1dvar$short[th.match],")"))
  n.th.vlist = length(th.vlist)
  #---~---

  #--- Proceed only if all variables exist in the output
  if (all(th.vlist %in% names(hlm1d))){
     #--- Title for legend
     cat0("   - ",th.thdesc,".")
     #---~---


     #--- Temporary data table. We convert the classes back to factor.
     thmelt       = hlm1d[,..th.vlist]
     thmelt$when  = chron(when)
     thmelt       = melt( data          = thmelt
                        , id.vars       = "when"
                        , variable.name = "idvar"
                        , measure.vars  = th.vlist
                        , value.name    = th.thnam
                        )#end melt
     thmelt       = data.table(thmelt)
     thmelt$idvar = factor(as.integer(thmelt$idvar))
     #---~---

       
     #--- Find colours for theme plot.
     th.funcol   = try(match.fun(th.vcols),silent=TRUE)
     if ("try-error" %in% is(th.funcol)){
       th.colour = c(unlist(strsplit(x=th.vcols,split="\\+")))
     }else{
       th.colour = th.funcol(n=n.th.vlist)
     }#end if ("try-error %in% is(th.funcol))
     #---~---
  
     
     
     #--- Initialise plot (decide whether to plot lines or stacks).
     if (th.stack){
        gg.now = ggplot(data=thmelt,aes_string(x="when",y=th.thnam,group="idvar",fill="idvar"))
        gg.now = gg.now + scale_fill_manual(name=character(0),labels=th.labels,values=th.colour)
        gg.now = gg.now + geom_area(position="stack",show.legend = TRUE)
     }else{
        gg.now = ggplot(data=thmelt,aes_string(x="when",y=th.thnam,group="idvar",colour="idvar"))
        gg.now = gg.now + scale_colour_manual(name=character(0),labels=th.labels,values=th.colour)
        gg.now = gg.now + geom_line(lwd=1.0,show.legend = TRUE)
     }#end if (f.stack)
     gg.now = gg.now + labs(title=simul.desc)
     gg.now = gg.now + scale_x_chron(format=gg.tfmt)
     gg.now = gg.now + xlab("Simulation time")
     gg.now = gg.now + ylab(desc.unit(desc=th.thdesc,unit=untab[[th.thunit]],twolines=TRUE))
     gg.now = gg.now + theme_grey( base_size      = gg.ptsz
                                 , base_family    = "Helvetica"
                                 , base_line_size = 0.5
                                 , base_rect_size = 0.5
                                 )#end theme_grey
     gg.now = gg.now + theme( legend.position   = "bottom"
                            , axis.text.x       = element_text( size   = gg.ptsz
                                                              , margin = unit(rep(0.35,times=4),"cm")
                                                              )#end element_text
                            , axis.text.y       = element_text( size   = gg.ptsz
                                                              , margin = unit(rep(0.35,times=4),"cm")
                                                              )#end element_text
                            , axis.ticks.length = unit(-0.25,"cm")
                            )#end theme
     #---~---
  
     #--- Save plot.
     for (d in sequence(ndevice)){
        th.output = paste0(th.thnam,"-tstheme-",simul,".",gg.device[d])
        dummy = ggsave( filename = th.output
                      , plot     = gg.now
                      , device   = gg.device[d]
                      , path     = tstheme.path
                      , width    = gg.width
                      , height   = gg.height
                      , units    = gg.units
                      , dpi      = gg.depth
                      )#end ggsave
     }#end for (o in sequence(nout))
     #---~---

     #--- Write plot settings to the list.
     gg.theme[[th.thnam]] = gg.now
     #---~---
    
  }#end if (all(th.vlist %in% names(hlm1d))))    
  #---~---
}#end for (a in age.loop)


#--- If sought, plot images on screen
if (gg.screen) gg.theme
#---~---

```
